---
title: "Taxonomy_RUMEN"
author: "Chiara Gini"
date: "21/02/2022"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library("DT")
library("knitr")
library("tidyr")
library("broom")
library("vegan")
library("ggpubr")
library("ggplot2")
library("reshape2")
library("tidyverse")
library("data.table")
library("ggrepel")
```

#COMPARISON BETWEEN TREATMENS 

## 16S sequencing data - cow rumen samples

```{r data, echo=FALSE}
metadata <- fread("~/Results/RUMEN/mapping_file_rumen.csv")
names(metadata)[1] <- "sample"

metadata$treatment[ which(metadata$treatment == "no treatment (ruminal liquid + diet)")] <- "Control"
metadata$treatment[ which(metadata$treatment == "AE1")] <- "EO1"
metadata$treatment[ which(metadata$treatment == "AE sintético 1")] <- "Synthetic EO1"

meta_subset <- filter(metadata, treatment !="ruminal liquid")
meta_subset %>%
  group_by(treatment) %>%
  dplyr::summarise(N=n())
```

```{r taxonomy, echo=FALSE}
otu <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L2.txt", header = TRUE, skip = 1)
otu <- gather(otu, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)
otu$treatment = metadata$treatment[match(otu$sample,metadata$sample)]
otu <- filter(otu, sample %in% meta_subset$sample)

## relative abundances
metadata_cols = names(meta_subset)[c(1,5)]
M <- select(otu,-all_of(metadata_cols))
M <- M/rowSums(M)
M <- bind_cols(select(otu, all_of(metadata_cols)),M)

M <- subset(M, treatment !="ruminal liquid")
```

```{r, echo=FALSE, fig.height=12, fig.width=12}
otu = otu %>% relocate (treatment, .after = sample)
metadata_cols = names(otu)[c(1,2)]

otu <- subset(otu, treatment !="ruminal liquid")

mO <- reshape2::melt(otu,id.vars = metadata_cols, value.name = "counts", variable.name = "phylum")
mO$phylum <- gsub("^.*;","",mO$phylum)
mO[mO$phylum == "2","phylum"] <- "WPS.2"
mO$sample <- factor(mO$sample)
mO <- mO %>%
  arrange(treatment)
#mO$Cow_ID <- paste(mO$cow, mO$ref)
```

```{r, echo=FALSE, fig.height=12, fig.width=12}
m1 <- reshape2::melt(M,id.vars = metadata_cols, value.name = "counts", variable.name = "phylum")
m1$phylum <- gsub("^.*;","",m1$phylum)
m1[m1$phylum == "2","phylum"] <- "WPS.2"
m1$sample <- factor(m1$sample)
m1 <- m1 %>%
  arrange(treatment)

# m1$sample <- paste(as.character(m1$sample),as.character(m1$ ),sep="-")
ordered_samples <- metadata %>%
  arrange(treatment) %>%
  select(treatment,sample)

m1$sample <- factor(m1$sample, levels = ordered_samples$sample)

m1$treatment <- factor(m1$treatment)
```

### Treatments 

First, let's look at phyla:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
D <- m1 %>%
  group_by(phylum,treatment) %>%
  summarise(N=n(), avg_abund = round(mean(counts),4), std = round(sd(counts),3)) %>%
  filter(avg_abund>0.005) %>%
  arrange(phylum,avg_abund)

phylum <- D %>%
  group_by(phylum) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  select(phylum,s)

D$phylum <- factor(D$phylum, levels = rev(phylum$phylum[order(phylum$s)]))
D <- D %>%
  arrange(treatment,phylum)

kable(D)
```

```{r, echo=FALSE}
p <- ggplot(mO, aes(x=phylum,y=as.factor(sample),fill=counts))
p <- p + geom_tile()
p <- p + facet_wrap(~treatment, scales = "free_y")
p <- p + ylab("sample")
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1),axis.text.y = element_text(size=9))
p <- p + scale_fill_gradient2(low="red",high = "blue")
p
#ggsave (p, height = 10, width = 20, filename = "taxonomy_count_rumen.png")
```

###boxplot abundances per treatment 

##Phyla

```{r, echo=FALSE}
p <- ggplot(D, aes(x=factor(1), y=avg_abund, fill=phylum)) + geom_bar(width=1,stat="identity")
p <- p + facet_grid(N~treatment)
#p <- p + coord_polar(theta='y', start=0) #activate this if you want a piechart
#p <- p + scale_fill_brewer(palette = "Set3") 
p <- p + geom_label_repel(aes(label = paste0(avg_abund*100, "%")), position=position_stack (vjust=0.5), size=4, show.legend = F)
p <- p + guides(fill = guide_legend(title = "Phyla")) 
p <- p + xlab("Relative abundances") + ylab("")
my_palette = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2"), length(unique(D$phylum)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))

print(p)
#ggsave(p, height = 20, width = 10, filename ="taxonomy_phylum_boxplot_rumen.png")
```
```{r}
fit = aov(avg_abund ~ treatment, data = D) ## linear model value = mu + treatment + e (y = mu + x + e)
TukeyHSD(fit, "treatment", ordered = TRUE)
```

### Treatments, all taxonomic levels

```{r, echo=FALSE,message=FALSE,warning=FALSE}
D <- D %>%
  group_by(treatment) %>%
  filter(avg_abund > 0.01) %>%
  arrange(desc(avg_abund))

phylum <- D %>%
  group_by(phylum) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  select(phylum,s)

D$phylum <- factor(D$phylum, levels = rev(phylum$phylum[order(phylum$s)]))
D <- D %>%
  arrange(phylum)

D$taxa_level = rep("phylum", nrow(D))

otu2 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L2.txt", header = TRUE, skip = 1)
otu3 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L3.txt", header = TRUE, skip = 1)
otu4 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L4.txt", header = TRUE, skip = 1)
otu5 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L5.txt", header = TRUE, skip = 1)
otu6 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L6.txt", header = TRUE, skip = 1)
otu <- bind_rows(otu2,otu3,otu4,otu5,otu6)

otu <- gather(otu, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)

otu$treatment = metadata$treatment[match(otu$sample,metadata$sample)]
otu <- filter(otu, sample %in% meta_subset$sample)

mO <- reshape2::melt(otu,id.vars = metadata_cols, value.name = "counts", variable.name = "taxa")

mO$level <- sapply(mO$taxa, function(x) length(strsplit(as.character(x),";")[[1]]) ) 

get_taxa <- function(x,n) {
  
  z <- strsplit(x, split = ";")
  res = lapply(z, function(x) x[n])
  return(unlist(res))
}

new_taxa <- c(NULL)
for (i in seq_along(mO$counts)) {
  
  arg1 = as.character(mO$taxa[i])
  arg2 = ifelse(mO$level[i] <= 6, mO$level[i], 6)
  
  rr = get_taxa(arg1,arg2)
  new_taxa <- c(new_taxa,rr)
}

mO$new_taxa <- new_taxa

table(mO$level)
neues <- c("phylum","class","order","family","genus")
altes <- seq(2,6)
mO$level <- neues[match(mO$level,altes)]

## relative abundances
mO <- mO %>%
  group_by(level,sample) %>%
  mutate(tot = sum(counts), rel_abundance = counts/tot)

D <- mO %>%
  group_by(level,new_taxa,treatment) %>%
  summarise(avg_abund = round(mean(rel_abundance),4), std = round(sd(rel_abundance),3))
```

A table of relative abundances (and standard deviations) per taxa and treatment:

```{r, tab2, echo=FALSE}
D <- na.omit(D)
D <- D %>%
  group_by(new_taxa,treatment) %>%
  filter(avg_abund > 0.01) %>%
  arrange(level,desc(avg_abund))

kable(D)
```

##Class

```{r}
Dclass <- D %>%
  filter(level=="class")

p <- ggplot(Dclass, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Class")) 
p <- p + geom_label_repel(aes(label = paste0(avg_abund*100, "%")), position=position_stack (vjust=0.5), size=4, show.legend = F)
my_palette = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple", "darkorange1","pink1", "darkorchid1", "seagreen2"), length(unique(Dclass$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))
print(p)
#ggsave(p, height = 8, width = 8, filename ="taxonomy_class_bargraph_rumen.png")
```
```{r}
fit = aov(avg_abund ~ treatment, data = Dclass) ## linear model value = mu + treatment + e (y = mu + x + e)
TukeyHSD(fit, "treatment", ordered = TRUE)
```

##Family

```{r}
Dfamily <- D %>%
  filter(level=="family")

p <- ggplot(Dfamily, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Family")) 
#p <- p + geom_label_repel(aes(label = paste0(avg_abund*100, "%")), position=position_stack (vjust=0.5), size=4, show.legend = F)
my_palette2 = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2", "darkmagenta", "burlywood1", "maroon", "darkred", "burlywood3", "chartreuse4", "coral", "peachpuff", "mediumorchid3", "yellow3", "turquoise4", "snow", "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(Dfamily$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))
p
#ggsave(p, height = 8, width = 8, filename ="taxonomy_family_bargraph_rumen.png")
```
```{r}
fit = aov(avg_abund ~ treatment, data = Dfamily) ## linear model value = mu + treatment + e (y = mu + x + e)
TukeyHSD(fit, "treatment", ordered = TRUE)
```

## Genera 

```{r}
Dgenera <- D %>%
  filter(level=="genus")

p <- ggplot(Dgenera, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Genera")) 
my_palette2 = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2", "darkmagenta", "burlywood1", "maroon", "darkred", "burlywood3", "chartreuse4", "coral", "peachpuff", "mediumorchid3", "yellow3", "turquoise4", "snow", "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(Dgenera$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=5),
               legend.title=element_text(size=5))
print(p)
ggsave(p, height = 8, width = 8, filename ="taxonomy_genera_bargraph_rumen.png")
```
```{r}
fit = aov(avg_abund ~ treatment, data = Dgenera) ## linear model value = mu + treatment + e (y = mu + x + e)
TukeyHSD(fit, "treatment", ordered = TRUE)
```

##Order

```{r}
Dorder <- D %>%
  filter(level=="order")

p <- ggplot(Dorder, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Order")) 
my_palette2 = get_palette(c("red","darkred", "burlywood3", "chartreuse4","gray","gold","blue","magenta","cyan","purple","green3","yellow","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2", "darkmagenta", "burlywood1", "maroon",  "coral", "peachpuff", "mediumorchid3", "yellow3", "turquoise4", "snow", "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(Dorder$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))
print(p)
#ggsave(p, height = 8, width = 8, filename ="taxonomy_order_bargraph_rumen.png")
```
```{r}
fit = aov(avg_abund ~ treatment, data = Dorder) ## linear model value = mu + treatment + e (y = mu + x + e)
TukeyHSD(fit, "treatment", ordered = TRUE)
```

Same results, but in wide-format (treatments on the horizontal axis): relative abundances (no standard deviations):

```{r, echo=FALSE}
D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))

D0 <- D %>%
  select(-c(std)) %>%
  spread(treatment,avg_abund)

# D0[is.na(D0)] <- ""

bg.picker <- function(z) {
  
  if(is.na(z)) {return(NULL)}
  if(is.numeric(z)) {
    
    if(z > 0.05 & z <= 0.10) {return("pink")}
    if(z > 0.10) {return("yellow")}
  }
}

library("formattable")
formattable(D0, list(
  
  `Control` = color_tile("white","green"),
  `Treated` = color_tile("white","deepskyblue")
  # `3 PR` = color_tile("white","yellow"),
  # `4 PR` = color_tile("white","pink"),
  # `5 PR` = color_tile("white","yellow"),
  # `6 PR` = formatter("span",
  #                 style = x ~ style(
  #                   "border-radius" = "4px",
  #                   "padding-right" = "4px",
  #                   "background-color" = sapply(x, bg.picker))
  #                 )
))

```

Finally,a bubble chart to visualize relative abundances per taxa and treatment:

```{r, echo=FALSE, fig.height = 14, fig.width = 10}
taxa <- D %>%
  group_by(level,new_taxa) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(level,s)) %>%
  select(level,new_taxa,s)

taxa <- taxa %>%
  group_by(level) %>%
  arrange(level,s)

p <- ggplot(D, aes(x = treatment, y = new_taxa))
p <- p + geom_point(aes(size = avg_abund, colour = treatment), alpha = 0.4)
p <- p + facet_grid(level~treatment, scales = "free", space = "free")
p <- p + scale_color_manual(values = c("green3","gold","blue","magenta","cyan","yellow", "green", "coral",  "#E7B800", "#FC4E07"))
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p <- p + theme(axis.text.x = element_text(size=8))
p <- p + theme_bw()
p <- p + xlab("treatments") + ylab("average abundances")
p
#ggsave (p, height = 13, width = 8, filename = "taxonomy_bubbleplot_rumen.png")
```

### Significance of treatments
 
The significance of differences between treatments (Control vs Treated with Essential Oil) was evaluated using a linear model (ANOVA) that included the effect of  s:
 
$$
counts_{ikj} = \mu +  _k + treatment_j + e_{ikj}
$$
# 
This way, the variability due to region is removed when assessing the effect of treatments.
 
```{r, echo=FALSE}
m1 <- mO %>%
  filter(!is.na(level)) %>%
  arrange(level,treatment)

D <- m1 %>%
  group_by(level, new_taxa) %>%
  do(tidy(anova(lm(counts ~  +treatment, data = .)))) %>%
  filter(term == "treatment")

D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>%
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))


#write.csv(D, "one.csv")
```


```{r, echo=FALSE}
DX <- D %>%
  filter(`p.value` <= 0.05) %>%
  select(c(level,new_taxa, `p.value`)) %>%
  arrange(level,`p.value`)

D0 <- mO %>%
  dplyr::group_by(level,new_taxa, treatment) %>%
  dplyr::summarise(avg_counts = mean(counts))

to_save = list(D,DX,D0)
#save(to_save, file = "taxonomy_ .RData")

load("taxonomy_ .RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

dd <- spread(D0, key = treatment, value = avg_counts)
temp <- inner_join(DX,dd, by = c("level" = "level", "new_taxa" = "new_taxa"))
#fwrite(temp, file = "~/~/Results/RUMEN/rumen_significant_otus.csv", col.names = TRUE, sep = ",")
print (dd)
```

```{r, echo=FALSE, warning=FALSE}
load("taxonomy_ .RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

D0 <- mutate(D0, avg_counts = avg_counts+1) %>% spread(key = treatment, value = avg_counts)

D1 <- DX %>%
  inner_join(D0, by = c("level" = "level", "new_taxa" = "new_taxa")) %>%
  mutate(p.value = -log10(p.value)) %>%
  gather(key = "treatment", value = "counts", -c(level,new_taxa, p.value))

D1$level <- factor(D1$level, levels = c("phylum","class","order","family","genus"))

D1 <- D1 %>% group_by(level) %>% mutate(tot = sum(counts), relab = counts/tot)

p <- ggplot(D1, aes(x = treatment, y = new_taxa))
p <- p + geom_tile(aes(fill = relab), colour = "white")
p <- p + facet_grid(level~treatment, scales = "free")
p <- p + scale_fill_gradient(low = "orange", high = "blue")
p <- p + theme(strip.text.y = element_text(size = 5), 
               strip.text.x = element_text(size = 6),
               # axis.text.y = element_text(size = 4),
               axis.text.x = element_text(size = 6),
               axis.title = element_text(size = 6))
p <- p + guides(fill="none") + theme(axis.title.y = element_blank(),
                                    axis.text.y = element_blank(),
                                    axis.ticks.y = element_blank())
# p

dd <- filter(D1, treatment == "EO1") %>% mutate(variable = "p-value")

q <- ggplot(dd, aes(x = factor(1), y = new_taxa, group=level))
q <- q + geom_tile(aes(fill = p.value), colour = "white")
q <- q + facet_grid(level~variable, space="free", scales = "free_y")
q <- q + scale_fill_gradient(low = "orange", high = "blue")
q <- q + theme(strip.text = element_text(size = 4), 
               strip.text.x = element_text(size = 6),
               axis.text.y = element_text(size = 5),
               axis.title = element_text(size = 6))
q <- q + guides(fill=FALSE) + theme(
  # axis.title.x = element_blank(),
  # axis.text.x=element_blank(),
  # axis.ticks.x=element_blank(),
  strip.text.y = element_blank(),
  # axis.text.x = element_blank()
  axis.text.x = element_text(size = 6)
)
q <- q + xlab("") 
# q

figure_final <- ggarrange(q, p, widths=c(0.5, 1), 
                          labels=c("A", "B"))

print(figure_final)
#ggsave(filename = "heatmap_rumen.png", plot = figure_final, device = "png", width = 8, height = 10)
```




#CHECK ON DIET EFFECTS ON RUMINAL FLUIDS

```{r data, echo=FALSE}
meta_subset_pre <- filter(metadata, treatment !="EO1" & treatment !="Carvacrol" & treatment !="Synthetic EO1" & treatment !="p-cymene" & treatment !="γ-terpinene")

meta_subset_pre %>%
  group_by(treatment) %>%
  dplyr::summarise(N=n())
```

```{r taxonomy, echo=FALSE}
otu <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L2.txt", header = TRUE, skip = 1)
otu <- gather(otu, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)
otu$treatment = metadata$treatment[match(otu$sample,metadata$sample)]
otu <- filter(otu, sample %in% meta_subset_pre$sample)

## relative abundances
metadata_cols = names(meta_subset_pre)[c(1,5)]
M <- select(otu,-all_of(metadata_cols))
M <- M/rowSums(M)
M <- bind_cols(select(otu, all_of(metadata_cols)),M)

M <- subset(M, treatment !="EO1" & treatment !="Carvacrol" & treatment !="Synthetic EO1" & treatment !="p-cymene" & treatment !="γ-terpinene")
```

```{r, echo=FALSE, fig.height=12, fig.width=12}
otu = otu %>% relocate (treatment, .after = sample)
metadata_cols = names(otu)[c(1,2)]

otu <- subset(otu, treatment !="EO1" & treatment !="Carvacrol" & treatment !="Synthetic EO1" & treatment !="p-cymene" & treatment !="γ-terpinene")

mO <- reshape2::melt(otu,id.vars = metadata_cols, value.name = "counts", variable.name = "phylum")
mO$phylum <- gsub("^.*;","",mO$phylum)
mO[mO$phylum == "2","phylum"] <- "WPS.2"
mO$sample <- factor(mO$sample)
mO <- mO %>%
  arrange(treatment)
#mO$Cow_ID <- paste(mO$cow, mO$ref)
```

```{r, echo=FALSE, fig.height=12, fig.width=12}
m1 <- reshape2::melt(M,id.vars = metadata_cols, value.name = "counts", variable.name = "phylum")
m1$phylum <- gsub("^.*;","",m1$phylum)
m1[m1$phylum == "2","phylum"] <- "WPS.2"
m1$sample <- factor(m1$sample)
m1 <- m1 %>%
  arrange(treatment)

# m1$sample <- paste(as.character(m1$sample),as.character(m1$ ),sep="-")
ordered_samples <- metadata %>%
  arrange(treatment) %>%
  select(treatment,sample)

m1$sample <- factor(m1$sample, levels = ordered_samples$sample)

m1$treatment <- factor(m1$treatment)
```

### Treatments 

First, let's look at phyla:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
D <- m1 %>%
  group_by(phylum,treatment) %>%
  summarise(N=n(), avg_abund = round(mean(counts),4), std = round(sd(counts),3)) %>%
  filter(avg_abund>0.005) %>%
  arrange(phylum,avg_abund)

phylum <- D %>%
  group_by(phylum) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  select(phylum,s)

D$phylum <- factor(D$phylum, levels = rev(phylum$phylum[order(phylum$s)]))
D <- D %>%
  arrange(treatment,phylum)

kable(D)
```

```{r, echo=FALSE}
p <- ggplot(mO, aes(x=phylum,y=as.factor(sample),fill=counts))
p <- p + geom_tile()
p <- p + facet_wrap(~treatment, scales = "free_y")
p <- p + ylab("sample")
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1),axis.text.y = element_text(size=9))
p <- p + scale_fill_gradient2(low="red",high = "blue")
p
#ggsave (p, height = 10, width = 20, filename = "taxonomy_count_rumen_pre.png")
```

###boxplot abundances per treatment 

##Phyla

```{r, echo=FALSE}
p <- ggplot(D, aes(x=factor(1), y=avg_abund, fill=phylum)) + geom_bar(width=1,stat="identity")
p <- p + facet_grid(~treatment)
#p <- p + coord_polar(theta='y', start=0) #activate this if you want a piechart
#p <- p + scale_fill_brewer(palette = "Set3") 
p <- p + geom_label_repel(aes(label = paste0(avg_abund*100, "%")), position=position_stack (vjust=0.5), size=4, show.legend = F)
p <- p + guides(fill = guide_legend(title = "Phyla")) 
p <- p + xlab("Relative abundances") + ylab("")
my_palette = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2"), length(unique(D$phylum)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))

print(p)
#ggsave(p, height = 20, width = 10, filename ="taxonomy_phylum_boxplot_rumen_pre.png")
```

### Treatments, all taxonomic levels

```{r, echo=FALSE,message=FALSE,warning=FALSE}
D <- D %>%
  group_by(treatment) %>%
  filter(avg_abund > 0.01) %>%
  arrange(desc(avg_abund))

phylum <- D %>%
  group_by(phylum) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  select(phylum,s)

D$phylum <- factor(D$phylum, levels = rev(phylum$phylum[order(phylum$s)]))
D <- D %>%
  arrange(phylum)

D$taxa_level = rep("phylum", nrow(D))

otu2 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L2.txt", header = TRUE, skip = 1)
otu3 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L3.txt", header = TRUE, skip = 1)
otu4 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L4.txt", header = TRUE, skip = 1)
otu5 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L5.txt", header = TRUE, skip = 1)
otu6 <- fread("~/Results/RUMEN/results/taxa_summary_abs/CSS_normalized_otu_table_L6.txt", header = TRUE, skip = 1)
otu <- bind_rows(otu2,otu3,otu4,otu5,otu6)

otu <- gather(otu, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)

otu$treatment = metadata$treatment[match(otu$sample,metadata$sample)]
otu <- filter(otu, sample %in% meta_subset_pre$sample)

mO <- reshape2::melt(otu,id.vars = metadata_cols, value.name = "counts", variable.name = "taxa")

mO$level <- sapply(mO$taxa, function(x) length(strsplit(as.character(x),";")[[1]]) ) 

get_taxa <- function(x,n) {
  
  z <- strsplit(x, split = ";")
  res = lapply(z, function(x) x[n])
  return(unlist(res))
}

new_taxa <- c(NULL)
for (i in seq_along(mO$counts)) {
  
  arg1 = as.character(mO$taxa[i])
  arg2 = ifelse(mO$level[i] <= 6, mO$level[i], 6)
  
  rr = get_taxa(arg1,arg2)
  new_taxa <- c(new_taxa,rr)
}

mO$new_taxa <- new_taxa

table(mO$level)
neues <- c("phylum","class","order","family","genus")
altes <- seq(2,6)
mO$level <- neues[match(mO$level,altes)]

## relative abundances
mO <- mO %>%
  group_by(level,sample) %>%
  mutate(tot = sum(counts), rel_abundance = counts/tot)

D <- mO %>%
  group_by(level,new_taxa,treatment) %>%
  summarise(avg_abund = round(mean(rel_abundance),4), std = round(sd(rel_abundance),3))
```

A table of relative abundances (and standard deviations) per taxa and treatment:

```{r, tab2, echo=FALSE}
D <- na.omit(D)
D <- D %>%
  group_by(new_taxa,treatment) %>%
  filter(avg_abund > 0.01) %>%
  arrange(level,desc(avg_abund))

kable(D)
```

##Class

```{r}
Dclass <- D %>%
  filter(level=="class")

p <- ggplot(Dclass, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Class")) 
p <- p + geom_label_repel(aes(label = paste0(avg_abund*100, "%")), position=position_stack (vjust=0.5), size=4, show.legend = F)
my_palette = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple", "darkorange1","pink1", "darkorchid1", "seagreen2"), length(unique(Dclass$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))
print(p)
#ggsave(p, height = 8, width = 8, filename ="taxonomy_class_bargraph_rumen_pre.png")
```

##Family

```{r}
Dfamily <- D %>%
  filter(level=="family")

p <- ggplot(Dfamily, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Family")) 
#p <- p + geom_label_repel(aes(label = paste0(avg_abund*100, "%")), position=position_stack (vjust=0.5), size=4, show.legend = F)
my_palette2 = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2", "darkmagenta", "burlywood1", "maroon", "darkred", "burlywood3", "chartreuse4", "coral", "peachpuff", "mediumorchid3", "yellow3", "turquoise4", "snow", "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(Dfamily$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))
p
#ggsave(p, height = 8, width = 8, filename ="taxonomy_family_bargraph_rumen_pre.png")
```

## Genera 

```{r}
Dgenera <- D %>%
  filter(level=="genus")

p <- ggplot(Dgenera, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Genera")) 
my_palette2 = get_palette(c("green3","gold","blue","magenta","cyan","yellow","red","gray","purple","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2", "darkmagenta", "burlywood1", "maroon", "darkred", "burlywood3", "chartreuse4", "coral", "peachpuff", "mediumorchid3", "yellow3", "turquoise4", "snow", "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(Dgenera$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=5),
               legend.title=element_text(size=5))
print(p)
#ggsave(p, height = 8, width = 8, filename ="taxonomy_genera_bargraph_rumen_pre.png")
```

```{r}
fit = aov(avg_abund ~ treatment, data = Dgenera) ## linear model value = mu + treatment + e (y = mu + x + e)
TukeyHSD(fit, "treatment", ordered = TRUE)
```
##Order

```{r}
Dorder <- D %>%
  filter(level=="order")

p <- ggplot(Dorder, aes(x=factor(1), y=avg_abund, fill=new_taxa)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid( ~treatment)
p <- p + guides(fill = guide_legend(title = "Order")) 
my_palette2 = get_palette(c("red","darkred", "burlywood3", "chartreuse4","gray","gold","blue","magenta","cyan","purple","green3","yellow","brown","black", "darkorange1","pink1", "darkorchid1", "seagreen2", "darkmagenta", "burlywood1", "maroon",  "coral", "peachpuff", "mediumorchid3", "yellow3", "turquoise4", "snow", "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(Dorder$new_taxa)))
p <- p + xlab("relative abundances") + ylab("percentages")
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))
print(p)
#ggsave(p, height = 8, width = 8, filename ="taxonomy_order_bargraph_rumen_pre.png")
```

Same results, but in wide-format (treatments on the horizontal axis): relative abundances (no standard deviations):

```{r, echo=FALSE}
D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))

D0 <- D %>%
  select(-c(std)) %>%
  spread(treatment,avg_abund)

# D0[is.na(D0)] <- ""

bg.picker <- function(z) {
  
  if(is.na(z)) {return(NULL)}
  if(is.numeric(z)) {
    
    if(z > 0.05 & z <= 0.10) {return("pink")}
    if(z > 0.10) {return("yellow")}
  }
}

library("formattable")
formattable(D0, list(
  
  `Control` = color_tile("white","green"),
  `Treated` = color_tile("white","deepskyblue")
  # `3 PR` = color_tile("white","yellow"),
  # `4 PR` = color_tile("white","pink"),
  # `5 PR` = color_tile("white","yellow"),
  # `6 PR` = formatter("span",
  #                 style = x ~ style(
  #                   "border-radius" = "4px",
  #                   "padding-right" = "4px",
  #                   "background-color" = sapply(x, bg.picker))
  #                 )
))

```

Finally,a bubble chart to visualize relative abundances per taxa and treatment:

```{r, echo=FALSE, fig.height = 14, fig.width = 10}
taxa <- D %>%
  group_by(level,new_taxa) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(level,s)) %>%
  select(level,new_taxa,s)

taxa <- taxa %>%
  group_by(level) %>%
  arrange(level,s)

p <- ggplot(D, aes(x = treatment, y = new_taxa))
p <- p + geom_point(aes(size = avg_abund, colour = treatment), alpha = 0.4)
p <- p + facet_grid(level~treatment, scales = "free", space = "free")
p <- p + scale_color_manual(values = c("green3","gold","blue","magenta","cyan","yellow", "green", "coral",  "#E7B800", "#FC4E07"))
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p <- p + theme(axis.text.x = element_text(size=8))
p <- p + theme_bw()
p <- p + xlab("treatments") + ylab("average abundances")
p
#ggsave (p, height = 13, width = 8, filename = "taxonomy_bubbleplot_rumen_pre.png")
```

### Significance of treatments
 
The significance of differences between treatments (Control vs Treated with Essential Oil) was evaluated using a linear model (ANOVA) that included the effect of  s:
 
$$
counts_{ikj} = \mu +  _k + treatment_j + e_{ikj}
$$
# 
This way, the variability due to region is removed when assessing the effect of treatments.
 
```{r, echo=FALSE}
m1 <- mO %>%
  filter(!is.na(level)) %>%
  arrange(level,treatment)

D <- m1 %>%
  group_by(level, new_taxa) %>%
  do(tidy(anova(lm(counts ~  +treatment, data = .)))) %>%
  filter(term == "treatment")

D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>%
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))


#write.csv(D, "one_pre.csv")
```


```{r, echo=FALSE}
DX <- D %>%
  filter(`p.value` <= 0.05) %>%
  select(c(level,new_taxa, `p.value`)) %>%
  arrange(level,`p.value`)

D0 <- mO %>%
  dplyr::group_by(level,new_taxa, treatment) %>%
  dplyr::summarise(avg_counts = mean(counts))

to_save = list(D,DX,D0)
save(to_save, file = "taxonomy_ .RData")

load("taxonomy_ .RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

dd <- spread(D0, key = treatment, value = avg_counts)
temp <- inner_join(DX,dd, by = c("level" = "level", "new_taxa" = "new_taxa"))
fwrite(temp, file = "~/Results/RUMEN/rumen_significant_otus_pre.csv", col.names = TRUE, sep = ",")
`print (dd)
```

```{r, echo=FALSE, warning=FALSE}
load("taxonomy_ .RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

D0 <- mutate(D0, avg_counts = avg_counts+1) %>% spread(key = treatment, value = avg_counts)

D1 <- DX %>%
  inner_join(D0, by = c("level" = "level", "new_taxa" = "new_taxa")) %>%
  mutate(p.value = -log10(p.value)) %>%
  gather(key = "treatment", value = "counts", -c(level,new_taxa, p.value))

D1$level <- factor(D1$level, levels = c("phylum","class","order","family","genus"))

D1 <- D1 %>% group_by(level) %>% mutate(tot = sum(counts), relab = counts/tot)

p <- ggplot(D1, aes(x = treatment, y = new_taxa))
p <- p + geom_tile(aes(fill = relab), colour = "white")
p <- p + facet_grid(level~treatment, scales = "free")
p <- p + scale_fill_gradient(low = "orange", high = "blue")
p <- p + theme(strip.text.y = element_text(size = 5), 
               strip.text.x = element_text(size = 6),
               # axis.text.y = element_text(size = 4),
               axis.text.x = element_text(size = 6),
               axis.title = element_text(size = 6))
p <- p + guides(fill="none") + theme(axis.title.y = element_blank(),
                                    axis.text.y = element_blank(),
                                    axis.ticks.y = element_blank())
# p

dd <- filter(D1, treatment == "EO1") %>% mutate(variable = "p-value")

q <- ggplot(dd, aes(x = factor(1), y = new_taxa, group=level))
q <- q + geom_tile(aes(fill = p.value), colour = "white")
q <- q + facet_grid(level~variable, space="free", scales = "free_y")
q <- q + scale_fill_gradient(low = "orange", high = "blue")
q <- q + theme(strip.text = element_text(size = 4), 
               strip.text.x = element_text(size = 6),
               axis.text.y = element_text(size = 5),
               axis.title = element_text(size = 6))
q <- q + guides(fill=FALSE) + theme(
  # axis.title.x = element_blank(),
  # axis.text.x=element_blank(),
  # axis.ticks.x=element_blank(),
  strip.text.y = element_blank(),
  # axis.text.x = element_blank()
  axis.text.x = element_text(size = 6)
)
q <- q + xlab("") 
# q

figure_final <- ggarrange(q, p, widths=c(0.5, 1), 
                          labels=c("A", "B"))

print(figure_final)
#ggsave(filename = "heatmap_rumen_pre.png", plot = figure_final, device = "png", width = 8, height = 10)
```
