---
title: "Beta_diversity_MILK"
author: "Chiara Gini"
date: "20/09/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library("knitr")
library("dplyr")
library("tidyr")
library("broom")
library("vegan")
library("ggplot2")
library("reshape2")
library("rmarkdown")
library("tidyverse")
library("ggfortify")
library("data.table")
library("ggforce")
library("plotly")
library("lattice")
```

## Beta diversity

From 16S rRNA-gene sequencing of cow milk microbiome samples:
- 4 timepoints for milk samples and two treatments (oil, treated/untreated)

Between-sample variability from 16S rRna-gene sequencing data.

```{r beta, echo=FALSE}
metadata <- fread("/home/mycelium/Results/Milkqua_skinswab_invivopublication/mapping_milkqua_skinswabs.csv")
names(metadata)[1] <- "sample"
names(metadata)[11] <- "Cow_ID"
metadata$timepoint[metadata$timepoint == "before_oil"] <- "T0"
# metadata$timepoint[metadata$timepoint == "after_oil"] <- "T0b"
metadata$timepoint[metadata$timepoint == "8"] <- "T7"
metadata$sample <-  gsub("sample-","sample.", as.character(metadata$sample))
metadata$sample2 <-  gsub("sample.","sample-", as.character(metadata$sample))
metadata <- subset(metadata, timepoint != "after_oil")
```

## Clustering

### PCA

```{r pressure, echo=FALSE, message=FALSE}
matrice= read.table("/home/mycelium/Results/Milkqua_skinswab_invivopublication/results/bray_curtis_distances.csv", row.names = 1, header=T, sep = ",")
matrice$sample <- rownames(matrice)
matrice$sample <- gsub("sample-", "sample.", as.character(matrice$sample))

vec <- filter(metadata, sample_type == "skinswab") %>% select(sample) %>% pull()
vex <- names(matrice) %in% vec
mat_skin = matrice[vex,vex]
# 
mat_skin$treatment <- as.character(metadata$treatment[match(row.names(mat_skin),metadata$sample2)])
mat_skin$timepoint <- as.character(metadata$timepoint[match(row.names(mat_skin),metadata$sample2)])

matx= data.matrix(select(mat_skin, -c(treatment,timepoint)))

## MDS
mds <- cmdscale(as.dist(matx))
```

### MDS

```{r}
mds <- cmdscale(as.dist(matx))
mds <- as.data.frame(mds)

mds$treatment <- metadata$treatment[match(rownames(mds), metadata$sample2)]
mds$timepoint <- metadata$timepoint[match(rownames(mds), metadata$sample2)]
mds$sample <- metadata$sample[match(rownames(mds), metadata$sample2)]
#mds <- mutate(mds, cow = as.factor(cow))

s <- ggplot(mds, aes(V1,V2)) + geom_point(aes(colour = timepoint, shape = treatment), size = 3) + stat_ellipse(aes(x=V1, y=V2,color=timepoint), type="norm")
s <- s + xlab("dim1") + ylab("dim2")
s

fname = file.path("~/Results/Milkqua_skinswab_invivopublication/results/beta/beta_SKINSWAB-2.png")
ggsave(filename = fname, plot = s, device = "png", dpi = 300, width = 30, height = 30)
```

## Significance of between-group distances

Significance values based on permuted analysis of variance (999 permutations), repeated 100 times.

```{r, label='permanova', echo=FALSE}
pv_treatment <- replicate(100, adonis2(matx ~ mat_skin$treatment, permutations = 999)$"Pr(>F)"[1], simplify = "vector")

pv_timepoint <- replicate(100, adonis2(matx ~ mat_skin$timepoint, permutations = 999)$"Pr(>F)"[1], simplify = "vector")

pv_treat_time <- replicate(100, adonis2(matx ~ mat_skin$treatment + mat_skin$timepoint + mat_skin$timepoint:mat_skin$treatment, permutations = 999)$"Pr(>F)"[1], simplify = "vector")


permanova <- function(mat,term,perms) {
  
  perm <- how(nperm = sample(perms,1))
  obj <- adonis2(mat ~ term, add = TRUE, parallel = TRUE, permutations = perm)
  return(obj$aov.tab$`Pr(>F)`[1])
}

pv_treat_time2 <- replicate(100,
  adonis2(matx ~ mat_skin$treatment + mat_skin$timepoint + mat_skin$timepoint:mat_skin$treatment, permutations = 999)$"Pr(>F)"[1],
  simplify = "vector"
)

```

- between treatments: p-value = **`r mean(pv_treatment)`** (std = `r sd(pv_treatment)`)
- between types: p-value = **`r mean(pv_timepoint)`** (std = `r sd(pv_timepoint)`)
- between treatments and type =  p-value = **`r mean(pv_treat_time)`** (std = `r sd(pv_treat_time)`)


Below the ANOVA table from the model:

$$
X = \mu + treatment + timepoint + treatment*type + e
$$
