---
title: "Alpha_diversity_stool"
author: "Chiara Gini"
date: "14/09/2022"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library("DT")
library("knitr")
library("dplyr")
library("tidyr")
library("broom")
library("vegan")
library("scales")
library("ggplot2")
#library("ggpubr")
# library("madbito")
library("reshape2")
library("gghighlight")
library("data.table")
library("rmarkdown")
library("ggrepel")
library("tibble")
library("ggplot2")
```

## Alpha diversity of stools samples

From 16S rRNA-gene sequencing of cow stools microbiome samples:
- 2 groups (Treated vs Control ), 2 timepoints.
Total: 32 samples.

Main alpha diversity indexes per group. 

```{r alpha, echo=FALSE}
metadata <- fread("/home/mycelium/Results/STOOLS/mapping_milkqua_stools.csv")
metadata <- tibble::rownames_to_column(metadata, "sample")


alpha <- read.table("~/Results/STOOLS/results/alpha.csv", header = TRUE, sep = ",")
alpha$sample.id <- row.names(alpha)
alpha$observed_species <- NULL

mAlpha <- reshape2::melt(alpha, id.vars = "sample.id", variable.name = "metric", value.name = "value")
mAlpha$timepoint <- metadata$timepoint[match(mAlpha$sample,metadata$sample)]
mAlpha$treatment <- metadata$treatment[match(mAlpha$sample,metadata$sample)]
mAlpha <- subset(mAlpha, timepoint == '3')
```

## Averages and plots

```{r pressure, echo=FALSE}
mAlpha$treatment <- factor(mAlpha$treatment, levels = c("control","EO"))

D <- mAlpha %>%
  group_by(metric,treatment) %>%
  summarize(N=n(),avg=round(mean(value),3)) %>%
  spread(key = metric, value = avg)

d_alpha <- D %>%
  gather(key = "metric", value = "value", -c(treatment,N)) %>%
  spread(key = "treatment", value = "value")

write_csv(d_alpha, "~/Results/STOOLS/results/alpha_stools_T3.csv")
kable(D)
```

### Boxplots of alpha diversity indexes

#### Treatment by region

```{r, echo=FALSE, fig.height=12, fig.width=12}
mAlpha <- mAlpha %>% group_by(metric) %>%
  mutate(scaled_value = rescale(value, to = c(0,100)))
  
p <- ggplot(mAlpha, aes(x=timepoint,y=scaled_value)) 
p <- p + geom_boxplot(aes(fill=treatment))
p <- p + scale_fill_manual(values=c("green3", "orange"))
p <- p + facet_wrap(~metric, scales = "free")
p <- p + xlab("Treatment")
p = p + theme(strip.text = element_text(size = 12),
              axis.text.x = element_text(angle = 90, size = 8))
p
```

## Testing for significance

### Original alpha-diversity index-values - effect of timepoint in stoolS samples

```{r test_alpha, echo=FALSE}
mAlpha$treatment = as.character(mAlpha$treatment)
mAlpha$treatment = factor(mAlpha$treatment, levels = c("control", "EO"))

D <- mAlpha %>%
  group_by(metric, timepoint) %>%
  do(tidy(lm(value ~ treatment, data = .))) %>%
  filter(term != "(Intercept)")

datatable(D, options = list(pageLength=100)) %>% 
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
D$timepoint <- factor(D$timepoint, levels=c("2", "3"))

p <- ggplot(data = D, mapping= aes(x=timepoint, y=p.value))
p <- p + geom_point(aes(color = metric, stroke = 1.5), position=position_jitter(h=0, w=0.25))
p <- p + geom_hline(yintercept=0.05, linetype="dashed", color = "red", size=0.5)
p <- p + geom_hline(yintercept=0.10, linetype="dashed", color = "darkorange", size=0.5)
#p <- p + scale_fill_manual(values=c("green3", "orange", "blue", "red", "yellow", "gray", "rose"))
p <- p + coord_trans(y="log2")
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_y_continuous(breaks=pretty_breaks(n=20)) + scale_y_continuous(breaks = c(0, 0.05, 0.10, 1))
#p <- p + gghighlight(p.value < 0.10, label_key = metric, use_direct_label = FALSE)
p <- p + theme(axis.text.x = element_text(angle = 90))
p

ggsave(filename = "~/Results/STOOLS/results/alpha_stools_T3.png", width = 4, height = 4, plot = p, device = "png")
```
```{r}
#temp <- filter(mAlpha, metric == "chao1")
# g = lm(value ~ treatment, data = temp)
# class(g)
fit = aov(value ~ treatment, data = mAlpha) ## linear model value = mu + treatment + e (y = mu + x + e)
TukeyHSD(fit, "treatment", ordered = TRUE)
```