---
title: "STOOLS_Taxonomy_OnlyControls"
author: "Chiara Gini"
date: "2022-11-24"
output: html_document
---

```{r setup, include=FALSE}
library("DT")
library("knitr")
library("tidyr")
library("broom")
library("vegan")
library("ggpubr")
library("ggplot2")
library("reshape2")
library("tidyverse")
library("data.table")
library("ggrepel")
library("tibble")
library("dplyr")
library("cowplot")
```
 
#COMPARISON BETWEEN TREATMENTS

```{r, echo=FALSE}
metadata <- fread("/home/mycelium/Results/STOOLS/mapping_milkqua_stools.csv")
metadata$sample <- paste0("sample-", metadata$MISEQ_ID)
metadata$timepoint[metadata$timepoint==2] <- "45 days old"
metadata$timepoint[metadata$timepoint==3] <- "13 months old"
metadata <- metadata[,c(7,8,10)]
metadata %>%
  group_by(timepoint,treatment) %>%
  dplyr::summarise(N=n())
metadata <- metadata[metadata$treatment == "control"] #CHANGE HERE
```

```{r, echo=FALSE}
otu <- fread("/home/mycelium/Results/STOOLS/results/otu_norm_CSS.csv", header = TRUE, sep = ",")
otu <- filter(otu, Family !="Mitochondria")
otu <- filter(otu, Class !="Chloroplast")
otu_tax =select(otu, -1, -34:-40)
otu_tax <- otu_tax/colSums(otu_tax)

otu_genus <- cbind(otu_tax, otu$Genus)
colnames(otu_genus)[33]<- "genera"
otu_genus <- otu_genus %>% group_by(genera) %>% summarise_all(funs(sum))

uncult <- slice(otu_genus, 1, 230:240)
uncult$genera <- "Uncultured or unknown"
uncult <- uncult %>%
  group_by(genera) %>%
  summarise(across(everything(), sum))

otu_genus <- otu_genus[-c(1, 230:240), ]
otu_genus <- rbind(otu_genus, uncult)

otu_genus <- as.data.frame(t(otu_genus))
colnames(otu_genus) <- otu_genus[1,]
otu_genus <- otu_genus[-c(1),]
otu_genus <- cbind(otu_genus, treatment=metadata$treatment, timepoint=metadata$timepoint, sample=metadata$sample)

metadata_cols = names(metadata)
```

###Relative abundances table

```{r relative abundances total table}
# otu_genus = subset(otu_genus, select = -c(timepoint, treatment, sample))

otu_tot <- otu_genus
taxa_names <- colnames(otu_tot)
#rownames(otu_tot) <- otu_tot[,241]
sample_names <- row.names(otu_tot)
#otu_tot <- transpose(otu_tot)
otu_tot <- as.data.frame(t(otu_tot))
colnames(otu_tot) <- sample_names
taxa_names <- as.data.frame(taxa_names)
otu_tot$`#OTU ID` <- cbind(taxa_names$taxa_names)
otu_tot <- otu_tot %>% dplyr::select(`#OTU ID`, everything())

otu_tot <- gather(otu_tot, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)

mO <- reshape2::melt(otu_tot,id.vars = metadata_cols, value.name = "counts", variable.name = "taxa")

mO$level <- sapply(mO$taxa, function(x) length(strsplit(as.character(x),";")[[1]]) ) 

get_taxa <- function(x,n) {
  
  z <- strsplit(x, split = ";")
  res = lapply(z, function(x) x[n])
  return(unlist(res))
}

new_taxa <- c(NULL)
for (i in seq_along(mO$counts)) {
  
  arg1 = as.character(mO$taxa[i])
  arg2 = ifelse(mO$level[i] <= 6, mO$level[i], 6)
  
  rr = get_taxa(arg1,arg2)
  new_taxa <- c(new_taxa,rr)
}

mO$new_taxa <- new_taxa

table(mO$level)
neues <- c("genus") #,"species"
altes <- seq(1) #2,6 if we want to include species
mO$level <- neues[match(mO$level,altes)]

## relative abundances
mO$counts <- as.numeric(mO$counts)

mO <- mO %>%
  group_by(level,sample) %>%
  mutate(tot = sum(counts), rel_abundance = counts/tot)

D <- mO %>%
  group_by(level,new_taxa,treatment,timepoint) %>%
  summarise(avg_abund = round(mean(rel_abundance),4), std = round(sd(rel_abundance),3))
```

A table of relative abundances (and standard deviations) per taxa and treatment:

```{r, tab2, echo=FALSE}
D <- na.omit(D)
D <- D %>%
  group_by(new_taxa,timepoint,treatment) %>%
#  filter(avg_abund > 0.01) %>% #ONE FILTER HERE
  arrange(level,desc(avg_abund))

kable(D)
```

### Significance of treatments per each timepoint
 
The significance of differences between treatments (Control vs Treated with Essential Oil) was evaluated using a linear model (ANOVA) that included the effect of timepoints:
 
$$
counts_{ikj} = \mu + treatment_j + e_{ikj}
$$
This way, the variability due to region is removed when assessing the effect of treatments.
 
```{r, echo=FALSE}
m1 <- mO %>%
  filter(!is.na(level)) %>%
  arrange(level,timepoint)

D <- m1 %>%
  group_by(level, new_taxa) %>%
  do(tidy(anova(lm(counts ~ timepoint, data = .)))) %>%
  filter(term == "timepoint")

D$level  <- factor(D$level,levels = c("genus"))
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>%
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))

# write.csv(D, "~/Results/STOOLS/results/T2/taxonomy_T2.csv")
```

### Significance of treatments (per timepoint)

The significance of differences between treatments (Control vs Treated) was evaluated using a linear model (ANOVA) that included the effect of time-points:

$$
counts_{ij} = \mu + treatment_j + e_{ij}
$$

this way, the variability due to region is removed when assessing the effect of treatments.

```{r, echo=FALSE}
m1 <- mO %>%
  arrange(level,timepoint)
save(m1, file = "~/Results/STOOLS/results/T2/Single_genera_signif_T2.RData")


dd_counts <- mO %>%
  group_by(level, new_taxa,timepoint,treatment) %>%
  # summarise(avg = mean(counts)) %>%
  summarise(avg = mean(rel_abundance)) %>%
  spread(key = "timepoint", value = "avg")

group_by(dd_counts, level) #%>% summarise(tot = sum(control))

D <- m1 %>%
  group_by(level, new_taxa, treatment) %>%
  do(tidy(anova(lm(counts ~ timepoint, data = .)))) %>%
  filter(term == "timepoint")

D$level  <- factor(D$level,levels = c("genus")) #, "species"
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>% 
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))
```

```{r, echo=FALSE}
DX <- D %>%
  filter(`p.value` <= 0.05) %>% #ONE FILTER HERE
  dplyr::select(c(level,new_taxa, `p.value`)) %>%
  arrange(level,`p.value`)

D0 <- mO %>%
  dplyr::group_by(level,new_taxa, timepoint,treatment) %>%
  dplyr::summarise(avg_counts = mean(counts))

to_save = list(D,DX,D0)
save(to_save, file = "~/Results/STOOLS/results/T2/taxonomy_stools_T2.RData")

load("~/Results/STOOLS/results/T2/taxonomy_stools_T2.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

dd <- spread(D0, key = timepoint, value = avg_counts)
temp <- inner_join(DX,dd, by = c("level" = "level", "new_taxa" = "new_taxa"))
fwrite(temp, file = "~/Results/STOOLS/results/T2/significant_otus_T2.csv", col.names = TRUE, sep = ",")
print (dd)

# temp$new_taxa <- paste(temp$new_taxa,", Timepoint:", temp$timepoint)
# temp <- temp %>% filter(level == "genus")
```


```{r}
Dna <- filter(temp, level %in% c("genus"))
## remember here to comment the taxonomic level you are using

Dna$T1vsT2 <- (Dna$`45 days old`-Dna$`13 months old`)
Dna$`13 months old` <-NULL
Dna$`45 days old` <-NULL

write.csv(Dna, "~/Results/STOOLS/results/taxonomy_genera.csv")

# Dna$timepoint = factor(Dna$timepoint, levels = c( "45 days old", "13 months old"))

q <- ggplot(Dna, aes(x = factor(1), y = new_taxa, height=0.95))
q <- q + geom_tile(aes(fill = p.value), colour = "white")
q <- q + scale_fill_gradient(low = "red", high = "green", limits=c(0,0.05)) #, guide = "none"
q <- q + xlab("pvalue") + ylab("Genera")
q <- q + theme(axis.text.x = element_text(size = 13),
               axis.text.y = element_text(size = 13),
             axis.ticks = element_blank(), 
             legend.position = "left",
             # axis.title.x = element_blank(),
             axis.title.y = element_text(size=14))
q

fg <- ggplot(Dna, aes(x=new_taxa, y=T1vsT2)) + geom_bar(width=0.95,stat="identity")
# fg <- fg + coord_flip() 
# + scale_fill_manual(values = "#E69F00", #0072B2",) #3
fg <- fg + coord_flip() + scale_fill_manual(values = c("#0072B2", "orange")) #2
fg <- fg + ylab("Behaviour Treated vs. Control") + xlab("Taxa") + ylim(-0.0314, 0.1094)
fg <- fg + theme(axis.text.x = element_text(size = 11),
                 axis.text.y = element_blank(),
                 axis.ticks = element_blank(),
                 axis.title.x = element_text(size=14),
               axis.title.y = element_blank())
fg

FF <- ggarrange(q, fg, widths=c(0.5,1), heights=c(1,1), labels = c("A","B","C","D"), legend = "left")
print(FF)
ggsave(filename = "~/Results/STOOLS/results/taxonomy_stools_OnlyControls.png", plot = FF, device = "png", width = 20, height = 20)
```
