---
title: "Milkqua_milk_invivopublication"
author: "Chiara Gini"
date: "2023-01-18"
output: html_document
---

```{r setup, include=FALSE}
library("DT")
library("knitr")
library("tidyr")
library("broom")
library("vegan")
library("ggpubr")
library("ggplot2")
library("reshape2")
library("tidyverse")
library("data.table")
library("ggrepel")
library("tibble")
library("dplyr")
library("cowplot")
library("ggpubr")
library("ggpubr")
```
 
#Metadata preparation

```{r , echo=FALSE}
basedir="~/Results"
prjdir = "Milkqua_milk_invivopublication"
outdir = "results"

metadata <- fread("/home/mycelium/Results/Milkqua_milk_invivopublication/mapping_milkqua_milk.csv")
names(metadata)[1] <- "sample"
names(metadata)[7] <- "type"

metadata$Cow_ID <- paste(metadata$cow, metadata$ref)
metadata$timepoint[metadata$timepoint == "0"] <- "T0"
metadata$timepoint[metadata$timepoint == "8"] <- "T7"
metadata$timepoint[metadata$timepoint == "21"] <- "T21"
metadata$timepoint[metadata$timepoint == "28"] <- "T28"
metadata$sample <- sub("^", "sample-", metadata$sample)
metadata <- metadata [order(metadata$sample),]

meta <- metadata [,c(1,11,12)]
meta_cols = names(meta)
metadata <- metadata[!(metadata$sample=="sample-80"),] #discarded due to low quality

otu <- fread("/home/mycelium/Results/Milkqua_milk_invivopublication/results/otu_norm_CSS.csv", header = TRUE, sep = ",")
otu <- filter(otu, Family !="Mitochondria")
otu <- filter(otu, Class !="Chloroplast")
otu <- filter(otu, Order !="Chloroplast")
```

#core microbiota phylum level

```{r}
otu_phylum=select(otu, 2:64,66)
otu_phylum <- otu_phylum %>% group_by(Phylum) %>% summarise(across(everything(),sum))
otu_box <- gather(otu_phylum, key = "sample", value ="counts", -Phylum) %>% spread(key = Phylum, value = counts)

## relative abundances
metadata_cols = names(metadata)[1]
M <- dplyr::select(otu_box,-all_of(metadata_cols))
M <- M/rowSums(M)
M <- bind_cols(dplyr::select(otu_box, all_of(metadata_cols)),M)

## plot of genus abundance
mm <- gather(M, key = "phylum", value = "abundance", -c(sample))
phyls = group_by(mm, phylum) %>% summarise(avg = mean(abundance)) %>% arrange(desc(avg))
oldc <- phyls$phylum[phyls$avg < 0.01]
newc <- rep("Lower than 1%", length(oldc))
vec <- newc[match(mm$phylum,oldc)]
mm$phylum <- ifelse(mm$phylum %in% oldc, "Lower than 1%", as.character(mm$phylum))
mm$phylum <- factor(mm$phylum, levels = c(phyls$phylum[1: (length(phyls$phylum) - length(oldc))],"Lower than 1%"))
```

```{r}
require('RColorBrewer')
mycolors = c(brewer.pal(name="Paired", n = 11), brewer.pal(name="Paired", n = 12))
mycolors2 = c(brewer.pal(name="Set2", n = 11), brewer.pal(name="Paired", n = 12))

p <- ggboxplot(mm, "phylum", "abundance", color = "phylum", legend = "none", palette = mycolors2)
p <- p + rotate_x_text(90) + font("xy.text", size=13)
p

A <- otu_phylum[,-1] > 0
vec <- rowSums(A)/ncol(A) > 0.99
A <- otu_phylum[vec,]
vec <- !grepl("uncultured", A$Phylum)
A <- A[vec,]
A$avg <- rowMeans(A[,-1])
A <- arrange(A, desc(avg)) %>% rename(taxon = Phylum)

## write out table of the core microbiota
ffname = file.path(basedir, prjdir, outdir, "taxonomy/core_microbiota_phylum.csv")
select(A, c(taxon,avg)) %>% rename(avg_normalised_counts = avg) %>% fwrite(ffname, sep = ",")

oldc <- A$taxon[A$avg < 25]
newc <- rep("Lower than 1%", length(oldc))
vec <- newc[match(A$taxon,oldc)]
A$taxon <- ifelse(A$taxon %in% oldc, "Lower than 1%", as.character(A$taxon))
A$taxon <- gsub("group","",A$taxon)

A <- group_by(A, taxon) %>% summarise(avg = mean(avg)) %>% arrange(desc(avg))
names(A)[1] <- "Phyla"

labs <- round(100*(A$avg/sum(A$avg)),2)
labs=paste0(labs, "%")

A$pct <- A$avg/sum(A$avg)
A$pctper <- round(100*(A$pct),2)
A$pctper = paste0(A$pctper, "%")
A$Phyla <- paste0(A$Phyla, " ", "(",A$pctper,")")
A <- A %>%
  arrange(desc(pct))
A$Phyla <- factor(A$Phyla)

pphyla <- ggplot(A, aes(x=factor(1), y=pct, fill=Phyla)) + geom_bar(width=1,stat="identity") 
pphyla <- pphyla + coord_polar(theta='y') + guides(fill = guide_legend(title = "Phyla")) 
pphyla <- pphyla + xlab("") + ylab("")
pphyla <- pphyla + scale_fill_manual(values = c("red", "green", "khaki", "steelblue1","maroon","lightblue","salmon","turquoise","violet","yellow","gray","tomato","navy","pink","springgreen4","peru","sienna1","plum4","mediumblue","darkorange","brown4","gold","bisque4"))
pphyla <- pphyla + theme(text = element_text(size=13),
               # axis.text.x = element_text(size=20),
               # strip.text = element_text(size = 20),
               axis.text = element_blank(),
               axis.ticks = element_blank(),
               legend.text=element_text(size=13),
               legend.title=element_text(size=13))
pphyla <- pphyla + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
pphyla
# ggsave(filename = file.path(basedir, prjdir, outdir, "Figure1perc.png"), plot = p, device = "png", dpi = 250, width = 21, height = 13)
```

#core microbiota phylum level

```{r}
otu_phylum=select(otu, 2:64,66)
otu_phylum <- otu_phylum %>% group_by(Phylum) %>% summarise(across(everything(),sum))
otu_box <- gather(otu_phylum, key = "sample", value ="counts", -Phylum) %>% spread(key = Phylum, value = counts)

## relative abundances
metadata_cols = names(metadata)[1]
M <- dplyr::select(otu_box,-all_of(metadata_cols))
M <- M/rowSums(M)
M <- bind_cols(dplyr::select(otu_box, all_of(metadata_cols)),M)

## plot of genus abundance
mm <- gather(M, key = "phylum", value = "abundance", -c(sample))
phyls = group_by(mm, phylum) %>% summarise(avg = mean(abundance)) %>% arrange(desc(avg))
oldc <- phyls$phylum[phyls$avg < 0.01]
newc <- rep("Lower than 1%", length(oldc))
vec <- newc[match(mm$phylum,oldc)]
mm$phylum <- ifelse(mm$phylum %in% oldc, "Lower than 1%", as.character(mm$phylum))
mm$phylum <- factor(mm$phylum, levels = c(phyls$phylum[1: (length(phyls$phylum) - length(oldc))],"Lower than 1%"))
```

###Effects of the treatments and time points on milk microbiota at phyla level.

```{r}
otu_norm <- select(otu_phylum, -1)
otu_norm <- otu_norm/colSums(otu_norm)
rownames(otu_norm) <- otu_phylum$Phylum

B <- as.data.frame(t(otu_norm))
B <- cbind(B, treatment=metadata$treatment, timepoint=metadata$timepoint, sample=metadata$sample)

m1 <- reshape2::melt(B,id.vars = meta_cols, value.name = "counts", variable.name = "phylum")
m1 <- m1 %>%
 arrange(treatment) 
ordered_samples <- metadata %>%
  arrange(metadata$treatment) %>%
  dplyr::select(treatment,sample)

m1$sample <- factor(m1$sample, levels = ordered_samples$sample)
m1$counts <- as.numeric(m1$counts)
m1$treatment <- factor(m1$treatment)
 
D <- m1 %>%
   group_by(phylum, treatment, timepoint) %>%
   summarise(N=n(), avg_abund = round(mean(counts),4), std = round(sd(counts),3)) %>%
   filter(avg_abund>0.01) %>% #ONE FILTER HERE
   # filter(timepoint==2) %>% #ONE FILTER HERE
   arrange(phylum,avg_abund)

phylum <- D %>%
  group_by(phylum) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  dplyr::select(phylum,s)

D$phylum <- factor(D$phylum, levels = rev(phylum$phylum[order(phylum$s)]))
D <- D %>%
  arrange(phylum)

kable(D)
```

#boxplot abundances per treatment

```{r, echo=FALSE}
D$treatment = factor(D$treatment, levels = c("Control", "Treated"), ordered = TRUE)

p <- ggplot(D, aes(x=factor(1), y=avg_abund, fill=phylum)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid(treatment~timepoint)
p <- p + guides(fill = guide_legend(title = "Phyla"))
my_palette2 = get_palette(c("brown", "darkorange1","pink1","seagreen2", "darkmagenta", "burlywood1",  "burlywood3", "chartreuse4", "coral", "peachpuff","gray", "mediumorchid3", "yellow3", "turquoise4", "green3","gold","blue","magenta","cyan","yellow","red","purple", "darkorchid1",  "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(D$phylum)))
p <- p + xlab("Relative abundances") + ylab("Percentages")
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=13),
               axis.text.x = element_text(size=13),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 13),
               axis.text.y = element_blank(),
               axis.ticks = element_blank(),
               legend.text=element_text(size=13),
               legend.title=element_text(size=13), legend.position = "none")
boxplotP <- p
print(p)
ggsave(p, height = 5, width = 12, filename ="~/Results/Milkqua_milk_invivopublication/results/taxonomy/phylum_boxplot.png")
```

###Relative abundances table - phyla

```{r relative abundances total table}
sample_names <- colnames(otu_norm)
taxa_names <- otu_phylum$Phylum
taxa_names <- as.data.frame(taxa_names)
otu_norm$`#OTU ID` <- cbind(taxa_names$taxa_names)
otu_norm <- otu_norm %>% dplyr::select(`#OTU ID`, everything())
otu_norm <- gather(otu_norm, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)
otu_norm <- otu_norm[order(otu_norm$sample),]
otu_norm <- cbind(otu_norm, treatment=metadata$treatment, timepoint=metadata$timepoint, sample=metadata$sample)
mO <- reshape2::melt(otu_norm,id.vars = meta_cols, value.name = "counts", variable.name = "taxa")

## relative abundances
mO$counts <- as.numeric(mO$counts)

mO <- mO %>%
  group_by(sample) %>%
  mutate(tot = sum(counts), rel_abundance = counts/tot)

D <- mO %>%
  group_by(taxa,treatment,timepoint) %>%
  summarise(avg_abund = round(mean(rel_abundance),4), std = round(sd(rel_abundance),3))

D <- na.omit(D)
D <- D %>%
  group_by(taxa,timepoint,treatment) %>%
#  filter(avg_abund > 0.01) %>% #ONE FILTER HERE
  arrange(desc(avg_abund))

kable(D)
```

### Significance of treatments per each timepoint - phyla

```{r, echo=FALSE}
mO$level <- "phylum"

m1 <- mO %>%
  filter(!is.na(level)) %>%
  arrange(level,treatment)
save(m1, file = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/Phylum_signif.RData")

dd_counts <- mO %>%
  group_by(level, taxa,timepoint,treatment) %>%
  # summarise(avg = mean(counts)) %>%
  summarise(avg = mean(rel_abundance)) %>%
  spread(key = "treatment", value = "avg")

group_by(dd_counts, level) %>% summarise(tot = sum(Treated))

D <- m1 %>%
  group_by(level, taxa, timepoint) %>%
  do(tidy(anova(lm(counts ~ treatment, data = .)))) %>%
  filter(term == "treatment")

D$level  <- factor(D$level,levels = c("phylum")) #, "species"
D <- D %>%
  arrange(level,taxa)

DX <- D %>%
  filter(`p.value` <= 0.05) %>% #ONE FILTER HERE
  dplyr::select(c(level,taxa, timepoint, `p.value`)) %>%
  arrange(level,`p.value`)

D0 <- mO %>%
  dplyr::group_by(level,taxa, timepoint,treatment) %>%
  dplyr::summarise(avg_counts = mean(counts))

to_save = list(D,DX,D0)
save(to_save, file = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/taxonomy-phyl.RData")

load("~/Results/Milkqua_milk_invivopublication/results/taxonomy/taxonomy-phyl.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

dd <- spread(D0, key = treatment, value = avg_counts)
temp <- inner_join(DX,dd, by = c("level" = "level", "taxa" = "taxa", "timepoint"="timepoint"))
fwrite(temp, file = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/significant_otus-phyl.csv", col.names = TRUE, sep = ",")
print (dd)
```

```{r, echo=FALSE, warning=FALSE}
load("~/Results/Milkqua_milk_invivopublication/results/taxonomy/taxonomy-phyl.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

D0 <- mutate(D0, avg_counts = avg_counts+1) %>% spread(key = treatment, value = avg_counts)

D1 <- DX %>%
  inner_join(D0, by = c("level" = "level", "taxa" = "taxa", "timepoint"="timepoint")) %>%
  mutate(p.value = -log10(p.value)) %>%
  gather(key = "treatment", value = "counts", -c(level,taxa,timepoint, p.value))
D1$level <- factor(D1$level, levels = c("phylum")) #, "species"
D1 <- D1 %>% group_by(level) %>% mutate(tot = sum(counts), relab = counts/tot)
```

### Extracting estimated coefficients for "significant" OTUs

```{r}
res = data.frame("level"=NULL, "OTU"=NULL, "p_value"=NULL, "Control"=NULL,"Treated"=NULL)

DX <- DX %>% filter(level == "phylum")
m1 <- m1 %>% filter(level=="phylum")
# DX$taxa <- paste(DX$taxa,", Timepoint:", DX$timepoint)
# m1$taxa <- paste(m1$taxa,", Timepoint:", m1$timepoint)


for (name in DX$taxa) {
  
  print(paste("analysing OTU ", name))
  pval = as.numeric(DX[DX$taxa==name,"p.value"])
  level = DX[DX$taxa==name,"level"]
  tpoint = DX[DX$taxa==name, "timepoint"]
  
  ## estimating coefficients
  temp = filter(m1, taxa == name)
  temp$treatment <- factor(temp$treatment, levels = c("Control","Treated"))
  g = lm(counts ~ treatment, data = temp)
  
  ## extracting coefficients
  coefs = g$coefficients
  coefs = coefs[!grepl("(Intercept)", names(coefs))]
  names(coefs) = gsub("treatment","",names(coefs))
  coefs = as.data.frame(t(coefs))
  
  ## adding metadata
  coefs["level"] = level
  coefs["OTU"] = name
  coefs["p_value"] = pval
  coefs["timepoint"] = tpoint

  # saving results
  res = rbind.data.frame(res, coefs)
}

write.csv(res, "~/Results/Milkqua_milk_invivopublication/results/taxonomy/pvalues-phyl.csv")
```

```{r}
Dna <- filter(temp, level %in% c("phylum"))

Dna$timepoint = factor(Dna$timepoint, levels = c("T0","T7","T21","T28"))

q <- ggplot(res, aes(x = factor(1), y = OTU, height=0.95))
q <- q + geom_tile(aes(fill = p_value), colour = "white")
q <- q + scale_fill_gradient(low = "#1aff1a", high = "#4b0092", limits=c(0,0.05)) #, guide = "none"
q <- q + xlab("pvalue") + ylab("Genera")
q <- q + theme(axis.text.x = element_text(size = 13),
               axis.text.y = element_text(size = 13),
             axis.ticks = element_blank(), 
             legend.position = "left",
             axis.title.x = element_text(size=13),
             axis.title.y = element_text(size=13))
q

fg <- ggplot(res, aes(x=OTU, y=Treated, fill=timepoint)) + geom_bar(width=0.95,stat="identity")
# fg <- fg + coord_flip() 
# + scale_fill_manual(values = "#E69F00", #0072B2",) #3
fg <- fg + coord_flip() + scale_fill_manual(values = c("#FFC20A", "#0C7BDC", "#e66100", "#40b0A6" )) #colorblind safe 
fg <- fg + ylab("Behaviour Treated vs. Control") + xlab("Taxa") + ylim(-0.455,0.0014) 
fg <- fg + theme(axis.text.x = element_text(size = 13),
                 axis.text.y = element_blank(),
                 axis.ticks = element_blank(),
                 # legend.position = "bottom",
                 axis.title.x = element_text(size=13),
               axis.title.y = element_blank())
fg

FG <- ggarrange(q, fg, widths=c(0.35,0.5), heights=c(1,1))


write.csv(Dna, "~/Results/Milkqua_milk_invivopublication/results/taxonomy/phyla-behav.csv")

ggsave(filename = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/acc-milk.png", plot = FG, device = "png", width = 20, height = 1.35, dpi=300)

#FF <- ggarrange(p, q, fg, widths=c(0.75, 0.35,0.5), heights=c(1,1,1), labels = c("A","B", "C"), ncol = 3)
#print(FF)
#ggsave(filename = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/tot-phyla-milk.png", plot = FF, device = "png", width = 15, height = 6)
```

#core microbiota genus level

```{r}
otu_genus=select(otu, 2:64,70)
otu_genus <- otu_genus %>% group_by(Genus) %>% summarise(across(everything(),sum))

uncult <- slice(otu_genus, 1, 377:386)
uncult$Genus <- "Uncultured or unknown"
uncult <- uncult %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum))

otu_genus <- otu_genus[-c(1, 377:386), ]
otu_genus <- rbind(otu_genus, uncult)
```

```{r, echo=FALSE}
otu_box <- gather(otu_genus, key = "sample", value ="counts", -Genus) %>% spread(key = Genus, value = counts)

## relative abundances
metadata_cols = names(metadata)[1]
M <- dplyr::select(otu_box,-all_of(metadata_cols))
M <- M/rowSums(M)
M <- bind_cols(dplyr::select(otu_box, all_of(metadata_cols)),M)

## plot of genus abundance
mm <- gather(M, key = "genus", value = "abundance", -c(sample))
phyls = group_by(mm, genus) %>% summarise(avg = mean(abundance)) %>% arrange(desc(avg))
oldc <- phyls$genus[phyls$avg < 0.01]
newc <- rep("Lower than 1%", length(oldc))
vec <- newc[match(mm$genus,oldc)]
mm$genus <- ifelse(mm$genus %in% oldc, "Lower than 1%", as.character(mm$genus))

mm$genus <- factor(mm$genus, levels = c(phyls$genus[1: (length(phyls$genus) - length(oldc))],"Lower than 1%"))
```

```{r}
library("ggpubr")

require('RColorBrewer')
mycolors = c(brewer.pal(name="Paired", n = 11), brewer.pal(name="Paired", n = 12))
mycolors2 = c(brewer.pal(name="Set2", n = 11), brewer.pal(name="Paired", n = 12))

p <- ggboxplot(mm, "genus", "abundance", color = "genus", legend = "none", palette = mycolors2)
p <- p + rotate_x_text(90) + font("xy.text", size=14)
p

A <- otu_genus[,-1] > 0
vec <- rowSums(A)/ncol(A) > 0.99
A <- otu_genus[vec,]
vec <- !grepl("uncultured", A$Genus)
A <- A[vec,]
A$avg <- rowMeans(A[,-1])
A <- arrange(A, desc(avg)) %>% rename(taxon = Genus)

## write out table of the core microbiota
ffname = file.path(basedir, prjdir, outdir, "taxonomy/core_microbiota.csv")
select(A, c(taxon,avg)) %>% rename(avg_normalised_counts = avg) %>% fwrite(ffname, sep = ",")

oldc <- A$taxon[A$avg < 25]
newc <- rep("Lower than 1%", length(oldc))
vec <- newc[match(A$taxon,oldc)]
A$taxon <- ifelse(A$taxon %in% oldc, "Lower than 1%", as.character(A$taxon))
A$taxon <- gsub("group","",A$taxon)

A <- group_by(A, taxon) %>% summarise(avg = mean(avg)) %>% arrange(desc(avg))
A$taxon -> A$short_name
A$short_name <- substr(A$short_name,start = 1, stop = 30)
A$short_name <- factor(A$short_name, levels = A$short_name)
names(A)[3] <- "Genera"

labs <- round(100*(A$avg/sum(A$avg)),2)
labs=paste0(labs, "%")

A$pct <- A$avg/sum(A$avg)
A$pctper <- round(100*(A$pct),2)
A$pctper = paste0(A$pctper, "%")
A$Genera <- paste0(A$Genera, " ", "(",A$pctper,")")
A <- A %>%
  arrange(desc(pct))
A$Genera <- factor(A$Genera)

pgenera <- ggplot(A, aes(x=factor(1), y=pct, fill=Genera)) + geom_bar(width=1,stat="identity") 
pgenera <- pgenera + coord_polar(theta='y') + guides(fill = guide_legend(title = "Genera")) 
pgenera <- pgenera + xlab("") + ylab("")
pgenera <- pgenera + scale_fill_manual(values = c("red", "green", "khaki", "steelblue1","maroon","lightblue","salmon","turquoise","violet","yellow","gray","tomato","navy","pink","springgreen4","peru","sienna1","plum4","mediumblue","darkorange","brown4","gold","bisque4"))
pgenera <- pgenera + theme(text = element_text(size=20),
               # axis.text.x = element_text(size=20),
               # strip.text = element_text(size = 20),
               axis.text = element_blank(),
               axis.ticks = element_blank(),
               legend.text=element_text(size=20),
               legend.title=element_text(size=20))
pgenera <- pgenera + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
pgenera

g <- ggarrange(pphyla, pgenera, ncol = 2, labels = c("A","B"), heights= c(1,1), widths = c(0.5,1)) #heights = c(0.1,4))
g

ggsave(filename = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/piechart_core_microbiome.png", plot = p, device = "png", dpi = 250, width = 21, height = 13)

library("cowplot")
fname = file.path(basedir, prjdir, outdir,"taxonomy/Figure1.png")
png(filename = fname, width = 20, height = 8, units = "in", res = 300)
ggdraw() +
  draw_plot(p, x = 0, y = 0, width = 0.3, height = 0.9) +
  draw_plot(q, x = 0.14, y = 0, width = 1, height = 1) +
  draw_plot_label(label = c("A", "B"), size = 16,
                  x = c(0, 0.32), y = c(1, 1)) 
dev.off()
```

###Effects of the treatments and time points on milk microbiota at genera level.

```{r}
otu_norm <- select(otu_genus, -1)
otu_norm <- otu_norm/colSums(otu_norm)
rownames(otu_norm) <- otu_genus$Genus

B <- as.data.frame(t(otu_norm))
B <- cbind(B, treatment=metadata$treatment, timepoint=metadata$timepoint, sample=metadata$sample)
meta_cols = names(meta)

m1 <- reshape2::melt(B,id.vars = meta_cols, value.name = "counts", variable.name = "genus")
m1 <- m1 %>%
  arrange(treatment)

ordered_samples <- metadata %>%
  arrange(metadata$treatment) %>%
  dplyr::select(treatment,sample)

m1$sample <- factor(m1$sample, levels = ordered_samples$sample)
m1$counts <- as.numeric(m1$counts)
m1$treatment <- factor(m1$treatment)

D <- m1 %>%
  group_by(genus, treatment, timepoint) %>%
  summarise(N=n(), avg_abund = round(mean(counts),4), std = round(sd(counts),3)) %>%
  filter(avg_abund>0.01) %>% #ONE FILTER HERE
  # filter(timepoint==2) %>% #ONE FILTER HERE
  arrange(genus,avg_abund)

genus <- D %>%
  group_by(genus) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  dplyr::select(genus,s)

D$genus <- factor(D$genus, levels = rev(genus$genus[order(genus$s)]))
D <- D %>%
  arrange(genus)

kable(D)

write.csv(D, "~/Results/Milkqua_milk_invivopublication/results/taxonomy/core_microbiota_gen.csv")
```

#boxplot abundances per treatment 

```{r, echo=FALSE}
D$treatment = factor(D$treatment, levels = c("Control", "Treated"), ordered = TRUE)

p <- ggplot(D, aes(x=factor(1), y=avg_abund, fill=genus)) + geom_bar(width=1,stat="identity", position = "fill")
p <- p + facet_grid(treatment~timepoint)
p <- p + guides(fill = guide_legend(title = "Genera")) 
my_palette2 = get_palette(c("brown", "darkorange1","pink1","seagreen2", "darkmagenta", "burlywood1",  "burlywood3", "chartreuse4", "coral", "peachpuff","gray", "mediumorchid3", "yellow3", "turquoise4", "green3","gold","blue","magenta","cyan","yellow","red","purple", "darkorchid1",  "slategray3", "orangered", "olivedrab", "plum", "lightcyan", "gray38", "forestgreen", "firebrick1"), length(unique(D$genus)))
p <- p + xlab("Relative abundances") + ylab("Percentages") 
p <- p + scale_fill_manual(values = my_palette2)
p <- p + theme(text = element_text(size=13),
               axis.text.x = element_text(size=13),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 13),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=13),
               legend.title=element_text(size=13), legend.position = "none")
boxplotG <- p
print(p)
ggsave(p, height = 5, width = 12, filename ="~/Results/Milkqua_milk_invivopublication/results/taxonomy/genera_boxplot.png")
```

###Relative abundances table

```{r relative abundances total table}
otu_norm <- select(otu_genus, -1)
otu_norm <- otu_norm/colSums(otu_norm)
rownames(otu_norm) <- otu_genus$Genus
sample_names <- colnames(otu_norm)
taxa_names <- otu_genus$Genus
taxa_names <- as.data.frame(taxa_names)
otu_norm$`#OTU ID` <- cbind(taxa_names$taxa_names)
otu_norm <- otu_norm %>% dplyr::select(`#OTU ID`, everything())
otu_norm <- gather(otu_norm, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)
otu_norm <- otu_norm[order(otu_norm$sample),]
otu_norm <- cbind(otu_norm, treatment=metadata$treatment, timepoint=metadata$timepoint, sample=metadata$sample)
mO <- reshape2::melt(otu_norm,id.vars = meta_cols, value.name = "counts", variable.name = "taxa")

## relative abundances
mO$counts <- as.numeric(mO$counts)

mO <- mO %>%
  group_by(sample) %>%
  mutate(tot = sum(counts), rel_abundance = counts/tot)

D <- mO %>%
  group_by(taxa,treatment,timepoint) %>%
  summarise(avg_abund = round(mean(rel_abundance),4), std = round(sd(rel_abundance),3))
```

A table of relative abundances (and standard deviations) per taxa and treatment:

```{r, tab2, echo=FALSE}
D <- na.omit(D)
D <- D %>%
  group_by(taxa,timepoint,treatment) %>%
#  filter(avg_abund > 0.01) %>% #ONE FILTER HERE
  arrange(desc(avg_abund))

kable(D)
```

### Significance of treatments per each timepoint
 
The significance of differences between treatments (Control vs Treated with Essential Oil) was evaluated using a linear model (ANOVA) that included the effect of timepoints:
 
$$
counts_{ikj} = \mu + treatment_j + e_{ikj}
$$
This way, the variability due to region is removed when assessing the effect of treatments.
 
```{r, echo=FALSE}
mO$level <- "genus"

m1 <- mO %>%
  filter(!is.na(level)) %>%
  arrange(level,treatment)

m1 <- mO %>%
  arrange(level,treatment)
save(m1, file = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/Genus_signif.RData")

dd_counts <- mO %>%
  group_by(level, taxa,timepoint,treatment) %>%
  # summarise(avg = mean(counts)) %>%
  summarise(avg = mean(rel_abundance)) %>%
  spread(key = "treatment", value = "avg")

group_by(dd_counts, level) %>% summarise(tot = sum(Treated))

D <- m1 %>%
  group_by(level, taxa, timepoint) %>%
  do(tidy(anova(lm(counts ~ treatment, data = .)))) %>%
  filter(term == "treatment")

D$level  <- factor(D$level,levels = c("genus")) #, "species"
D <- D %>%
  arrange(level,taxa)

datatable(D, options = list(pageLength=100)) %>% 
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))
```

```{r, echo=FALSE}
DX <- D %>%
  filter(`p.value` <= 0.05) %>% #ONE FILTER HERE
  dplyr::select(c(level,taxa, timepoint, `p.value`)) %>%
  arrange(level,`p.value`)

D0 <- mO %>%
  dplyr::group_by(level,taxa, timepoint,treatment) %>%
  dplyr::summarise(avg_counts = mean(counts))

to_save = list(D,DX,D0)
save(to_save, file = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/taxonomy-gen.RData")

load("~/Results/Milkqua_milk_invivopublication/results/taxonomy/taxonomy-gen.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

dd <- spread(D0, key = treatment, value = avg_counts)
temp <- inner_join(DX,dd, by = c("level" = "level", "taxa" = "taxa", "timepoint"="timepoint"))
fwrite(temp, file = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/significant_otus-gen.csv", col.names = TRUE, sep = ",")
print (dd)
```

```{r, echo=FALSE, warning=FALSE}
load("~/Results/Milkqua_milk_invivopublication/results/taxonomy/taxonomy-gen.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

D0 <- mutate(D0, avg_counts = avg_counts+1) %>% spread(key = treatment, value = avg_counts)

D1 <- DX %>%
  inner_join(D0, by = c("level" = "level", "taxa" = "taxa", "timepoint"="timepoint")) %>%
  mutate(p.value = -log10(p.value)) %>%
  gather(key = "treatment", value = "counts", -c(level,taxa,timepoint, p.value))

D1$level <- factor(D1$level, levels = c("genus")) #, "species"

D1 <- D1 %>% group_by(level) %>% mutate(tot = sum(counts), relab = counts/tot)
```

### Extracting estimated coefficients for "significant" OTUs

```{r}
res = data.frame("level"=NULL, "OTU"=NULL, "p_value"=NULL, "Control"=NULL,"Treated"=NULL)

DX <- DX %>% filter(level == "genus")
m1 <- m1 %>% filter(level=="genus")
# DX$taxa <- paste(DX$taxa,", Timepoint:", DX$timepoint)
# m1$taxa <- paste(m1$taxa,", Timepoint:", m1$timepoint)


for (name in DX$taxa) {
  
  print(paste("analysing OTU ", name))
  pval = as.numeric(DX[DX$taxa==name,"p.value"])
  level = DX[DX$taxa==name,"level"]
  tpoint = DX[DX$taxa==name, "timepoint"]
  
  ## estimating coefficients
  temp2 = filter(m1, taxa == name)
  temp2$treatment <- factor(temp2$treatment, levels = c("Control","Treated"))
  g = lm(counts ~ treatment, data = temp2)
  
  ## extracting coefficients
  coefs = g$coefficients
  coefs = coefs[!grepl("(Intercept)", names(coefs))]
  names(coefs) = gsub("treatment","",names(coefs))
  coefs = as.data.frame(t(coefs))
  
  ## adding metadata
  coefs["level"] = level
  coefs["OTU"] = name
  coefs["p_value"] = pval
  coefs["timepoint"] = tpoint

  # saving results
  res = rbind.data.frame(res, coefs)
}

write.csv(res, "~/Results/Milkqua_milk_invivopublication/results/taxonomy/pvalues-gen.csv")
```

```{r}
Dna <- filter(temp, level %in% c("genus"))
Dna$controlvstreated <- (Dna$Control-Dna$Treated)
Dna$Control <-NULL
Dna$Treated <-NULL

Dna$timepoint = factor(Dna$timepoint, levels = c("T0","T7","T21","T28"))

q <- ggplot(Dna, aes(x = factor(1), y = taxa, height=0.95))
q <- q + geom_tile(aes(fill = p.value), colour = "white")
q <- q + scale_fill_gradient(low = "#1aff1a", high = "#4b0092", limits=c(0,0.05)) #, guide = "none"
q <- q + xlab("") + ylab("Genera")
q <- q + theme(axis.text.x = element_text(size = 13),
               axis.text.y = element_text(size = 13),
             axis.ticks = element_blank(), 
             legend.position = "left",
             axis.title.x = element_text(size=13),
             axis.title.y = element_text(size=13))
q

fg <- ggplot(Dna, aes(x=taxa, y=controlvstreated, fill=timepoint)) + geom_bar(width=0.95,stat="identity")
# fg <- fg + coord_flip() 
# + scale_fill_manual(values = "#E69F00", #0072B2",) #3
fg <- fg + coord_flip() + scale_fill_manual(values = c("#FFC20A", "#0C7BDC", "#e66100", "#40b0A6" )) #colorblind safe 
fg <- fg + ylab("") + xlab("Taxa") + ylim(-0.03, 0.03)
fg <- fg + theme(axis.text.x = element_text(size = 13),
                 axis.text.y = element_blank(),
                 axis.ticks = element_blank(),
                 # legend.position = "top",
                 axis.title.x = element_blank(),
               axis.title.y = element_blank())
fg

write.csv(Dna, "~/Results/Milkqua_milk_invivopublication/results/taxonomy/genera-behav.csv")

FF <- ggarrange(q, fg, widths=c(0.35,0.65), heights=c(1,1))
print(FF)

ggsave(filename = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/tot-genera-milk.png", device = "png", dpi = 300, width = 20, height = 6)

#FF <- ggarrange(p, q, NULL, fg, widths=c(0.5,0.025,1), heights=c(1,1,1), labels = c("A","","B"), label.x = -0.015 , nrow = 1)
#print(FF)
#ggsave(filename = "~/Results/Milkqua_milk_invivopublication/results/taxonomy/taxonomy_barplot_pval-gen.png", plot = FF, device = "png", width = 15, height = 5)
```

-----------------------------------------------------------------------------------------------------------------------------------------------------------------

#Check the pvalue of everytreatment against control, genus by genus

ADD A FOR LOOP HERE ASAP

```{r}
library("broom")
library("tidyverse")

load("~/Results/Milkqua_milk_invivopublication/m1.Rdata")
m1 = to_save[[1]]
unique(m1$treatment)
m1$treatment = factor(m1$treatment, levels = c("Control","Carvacrol","G-terpinene", "NEO", "P-cymene","SEO"))

temp = filter(m1, level == "genus", taxa == "Lachnospiraceae UCG-010") ##specify genus here
fit <- lm(counts ~ treatment, data = temp)
anova(fit)
tt = summary(fit)
names(tt)

tt = tidy(fit)
filter(tt, term != "(Intercept)") %>% pull(p.value)
```

```{r}
load("~/Results/Milkqua_milk_invivopublication/m1.Rdata")
m1 = to_save[[1]]
unique(m1$treatment)
m1$treatment = factor(m1$treatment, levels = c("Control","Carvacrol","G-terpinene", "NEO", "P-cymene","SEO"))

temp = filter(m1, level == "genus", taxa == "Synergistes") ##specify genus here


temp1= filter(temp, treatment == "SEO", level== "genus")
mean(temp1$counts)
mean(temp1$rel_abundance)
```



