---
title: "Taxonomy_SHUVOmilk"
author: "Chiara Gini"
date: "11/10/2022"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library("DT")
library("knitr")
library("tidyr")
library("broom")
library("vegan")
library("ggpubr")
library("ggplot2")
library("reshape2")
library("tidyverse")
library("data.table")
library("ggrepel")
```
 
#OVERALL COMPARISON 

```{r, echo=FALSE}
#project_folder = "~/Documents/MILKQUA"
metadata <- fread("/home/mycelium/Results/SHUVO-MILK/mapping_shuvo_milk.csv")
names(metadata)[1] <- "sample"
metadata = filter(metadata, sample != "sample-46")
metadata <- select(metadata, -2:-6, -8:-13, -15:-21)

meta_subset <- metadata %>%
  group_by(sample_type) %>%
  dplyr::summarise(N=n())

kable(meta_subset)
```

```{r, echo=FALSE}
otu <- fread("/home/mycelium/Results/SHUVO-MILK/results/otu_norm_CSS.csv", header = TRUE)
otu = filter(otu, Family != "Mitochondria") ##filter to remove plant contamination
otu = filter(otu, Order != "Chloroplast") ##filter to remove plant contamination

otu_tax =select(otu, -1, -47:-53)
otu_tax <- otu_tax/colSums(otu_tax) ##realtive abundance
otu_genus <- cbind(otu_tax, otu$Genus)
colnames(otu_genus)[46]<- "Genus"
otu_genus <- otu_genus %>% group_by(Genus) %>% summarise_all(funs(sum))

uncult <- slice(otu_genus, 1, 283:290)
uncult$Genus <- "Uncultured or unknown"
uncult <- uncult %>%
  group_by(Genus) %>%
  summarise(across(everything(), sum))

otu_genus <- otu_genus[-c(1, 283:290), ]
otu_genus <- rbind(otu_genus, uncult)


otu_genus <- gather(otu_genus, key = "sample", value ="counts", -Genus) %>% spread(key = Genus, value = counts)
```

### Core microbiota as genera

```{r genera average abundances, echo=FALSE}
metadata_cols = names(metadata)

otu_genus <- otu_genus %>% inner_join(select(metadata, all_of(metadata_cols)), by="sample")

mO <- reshape2::melt(otu_genus,id.vars = metadata_cols, value.name = "counts", variable.name = "genus")
mO <- mO %>%
  arrange(sample_type)

ordered_samples <- metadata %>%
  arrange(metadata$sample_type) %>%
  dplyr::select(sample_type,sample)

mO$sample <- factor(mO$sample, levels = ordered_samples$sample)
mO$counts <- as.numeric(mO$counts)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
D <- mO %>%
  group_by(genus) %>%
  summarise(N=n(), avg_abund = round(mean(counts),4), std = round(sd(counts),3)) %>%
  filter(avg_abund>0.01) %>%
  arrange(genus,avg_abund)
genus <- D %>%
  group_by(genus) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  dplyr::select(genus,s)

D$genus <- factor(D$genus, levels = rev(genus$genus[order(genus$s)]))
D <- D %>%
  arrange(genus)

kable(D)

write.csv(D, "~/Results/SHUVO-MILK/results/core_microbiota.csv", sep = ",")
```
```{r, echo=FALSE}
D1 <- aggregate(cbind(D$avg_abund, D$std) ~ genus, data=D, FUN = sum)
names(D1)[2] <- "avg_abund"
names(D1)[3] <- "std"
kable(D1)

pgeneratot <- ggplot(D1, aes(x=factor(1), y=avg_abund, fill=genus)) + geom_bar(width=1,stat="identity")
pgeneratot <- pgeneratot + coord_polar(theta='y') #activate this if you want a piechart
pgeneratot <- pgeneratot + guides(fill = guide_legend(title = "Genera")) 
my_palette = get_palette(c("green3","yellow","blue","magenta","cyan","red","purple","brown", "darkorange1","pink1","darkorchid1", "seagreen2"), length(unique(D1$genus)))
pgeneratot <- pgeneratot + guides(fill = guide_legend(title = "genera")) 
pgeneratot <- pgeneratot + xlab("Relative abundances") + ylab("Percentages") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) + scale_fill_manual(values = my_palette)
pgeneratot <- pgeneratot + theme(text = element_text(size=10),
               axis.text.x = element_blank(),
               strip.text = element_text(size = 8),
              legend.position="bottom",
              axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))

print(pgeneratot)

ggsave(pgeneratot, height = 5, width = 15, filename ="~/Results/SHUVO-MILK/results/core_microbiota_piechart.png")
```

###genera for sample type

```{r}
m1 <- reshape2::melt(otu_genus,id.vars = metadata_cols, value.name = "counts", variable.name = "genus")
m1 <- m1 %>%
  arrange(sample_type)

ordered_samples <- metadata %>%
  arrange(metadata$sample_type) %>%
  dplyr::select(sample_type,sample)

m1$sample <- factor(m1$sample, levels = ordered_samples$sample)
m1$counts <- as.numeric(m1$counts)

D <- m1 %>%
  group_by(genus, sample_type) %>%
  summarise(N=n(), avg_abund = round(mean(counts),4), std = round(sd(counts),3)) %>%
  filter(avg_abund>0.01) %>%
  arrange(genus,avg_abund)

oldc <- D$genus[D$avg_abund < 0.01]
newc <- rep("Other", length(oldc))
vec <- newc[match(D$genus,oldc)]
D$genus <- ifelse(D$genus %in% oldc, "Other", as.character(D$genus))

genus <- D %>%
  group_by(genus) %>%
  summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  dplyr::select(genus,s)

D$genus <- factor(D$genus, levels = rev(genus$genus[order(genus$s)]))
D <- D %>%
  arrange(genus)

kable(D)
```

```{r plot genera, echo=FALSE}
D$sample_type <- factor(D$sample_type, levels = c("Healthy quarter", "Subclinical quarter", "Farm bulk milk"))

Pgenera <- ggplot(D, aes(x=factor(1), y=avg_abund, fill=genus)) + geom_bar(width=1,stat="identity", position = "fill")
Pgenera <- Pgenera + facet_grid(~sample_type)
#Pgenera <- Pgenera + coord_polar(theta='y', start=0) #activate this if you want a piechart
#Pgenera <- Pgenera + geom_label_repel(aes(label = paste0(avg_abund*100, "%")))
Pgenera <- Pgenera + guides(fill = guide_legend(title = "Genera")) 
my_palette = get_palette(c("green3","yellow","blue","magenta","cyan","red","purple","brown", "darkorange1","pink1","darkorchid1", "seagreen2"), length(unique(D$genus)))
Pgenera <- Pgenera + xlab("relative abundances") + ylab("percentages")
Pgenera <- Pgenera + scale_fill_manual(values = my_palette) + theme( panel.background = element_blank())
Pgenera <- Pgenera + theme(text = element_text(size=10),
               axis.text.x = element_text(size=8),
               # axis.text.y = element_text(size=4),
               strip.text = element_text(size = 8),
                legend.position="bottom",
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))

print(Pgenera)
ggsave(Pgenera, height = 3, width = 12, filename ="~/Results/SHUVO-MILK/results/genus_boxplot.png")

D %>% group_by(genus) %>% summarise(sample_type=n())
D %>% group_by(sample_type) %>% summarise(genus=n())
```

###Relative abundances table

```{r relative abundances total table}
otu_tot <- otu_genus
taxa_names <- colnames(otu_tot)
sample_names <- row.names(otu_tot)
#otu_tot <- transpose(otu_tot)
otu_tot <- as.data.frame(t(otu_tot))
colnames(otu_tot) <- sample_names
taxa_names <- as.data.frame(taxa_names)
otu_tot$`#OTU ID` <- cbind(taxa_names$taxa_names)
otu_tot <- otu_tot %>% dplyr::select(`#OTU ID`, everything())

otu_tot <- gather(otu_tot, key = "sample", value ="counts", -`#OTU ID`) %>% spread(key = `#OTU ID`, value = counts)
mO <- reshape2::melt(otu_tot,id.vars = metadata_cols, value.name = "counts", variable.name = "taxa")
mO = filter(mO, taxa != "type") ##filter to remove type in farm bulk milk

mO$level <- sapply(mO$taxa, function(x) length(strsplit(as.character(x),";")[[1]]) ) 

get_taxa <- function(x,n) {
  
  z <- strsplit(x, split = ";")
  res = lapply(z, function(x) x[n])
  return(unlist(res))
}

new_taxa <- c(NULL)
for (i in seq_along(mO$counts)) {
  
  arg1 = as.character(mO$taxa[i])
  arg2 = ifelse(mO$level[i] <= 6, mO$level[i], 6)
  
  rr = get_taxa(arg1,arg2)
  new_taxa <- c(new_taxa,rr)
}

mO$new_taxa <- new_taxa

table(mO$level)
neues <- c("genus") #,"species"
altes <- seq(1) #2,6 if we want to include species
mO$level <- neues[match(mO$level,altes)]

## relative abundances
 mO$counts <- as.numeric(mO$counts)

mO <- mO %>%
  group_by(level,sample) %>%
  mutate(tot = sum(counts), rel_abundance = counts/tot)

D <- mO %>%
  group_by(level,new_taxa,sample_type) %>%
  summarise(avg_abund = round(mean(rel_abundance),4), std = round(sd(rel_abundance),3))
```

A table of relative abundances (and standard deviations) per taxa and sample_type:

```{r, tab2, echo=FALSE}
D <- na.omit(D)
D <- D %>%
  group_by(new_taxa,sample_type) %>%
#  filter(avg_abund > 0.01) %>%
  arrange(level,desc(avg_abund))

kable(D)
```


### Significance of sample_types per each timepoint
 
The significance of differences between sample_types (Control vs Treated with Essential Oil) was evaluated using a linear model (ANOVA) that included the effect of timepoints:
 
$$
counts_{ikj} = \mu + sample_type_j + e_{ikj}
$$
This way, the variability due to region is removed when assessing the effect of sample_types.
 
```{r, echo=FALSE}
m1 <- mO %>%
  filter(!is.na(level)) %>%
  arrange(level,sample_type)

D <- m1 %>%
  group_by(level, new_taxa) %>%
  do(tidy(anova(lm(counts ~ sample_type, data = .)))) %>%
  filter(term == "sample_type")

D$level  <- factor(D$level,levels = c("genus"))
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>%
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))

write.csv(D, "~/Results/SHUVO-MILK/results/taxonomy_one.csv")
```

### Significance of sample_types (per timepoint)

The significance of differences between sample_types (Control vs Treated) was evaluated using a linear model (ANOVA) that included the effect of time-points:

$$
counts_{ij} = \mu + sample_type_j + e_{ij}
$$

this way, the variability due to region is removed when assessing the effect of sample_types.

```{r, echo=FALSE}
m1 <- mO %>%
  arrange(level,sample_type)

dd_counts <- mO %>%
  group_by(level, new_taxa,sample_type) %>%
  # summarise(avg = mean(counts)) %>%
  summarise(avg = mean(rel_abundance)) %>%
  spread(key = "sample_type", value = "avg")

group_by(dd_counts, level) %>% summarise(tot = sum(`Healthy quarter`))

D <- m1 %>%
  group_by(level, new_taxa) %>%
  do(tidy(anova(lm(counts ~ sample_type, data = .)))) %>%
  filter(term == "sample_type")

D$level  <- factor(D$level,levels = c("genus")) #, "species"
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>% 
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))
```

```{r, echo=FALSE}
DX <- D %>%
  filter(`p.value` <= 0.05) %>%
  dplyr::select(c(level,new_taxa, `p.value`)) %>%
  arrange(level,`p.value`)

D0 <- mO %>%
  dplyr::group_by(level,new_taxa, sample_type) %>%
  dplyr::summarise(avg_counts = mean(counts))

to_save = list(D,DX,D0)
save(to_save, file = "~/Results/SHUVO-MILK/results/taxonomy_SHUVO-MILK_tot.RData")

load("~/Results/SHUVO-MILK/results/taxonomy_SHUVO-MILK_tot.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

dd <- spread(D0, key = sample_type, value = avg_counts)
temp <- inner_join(DX,dd, by = c("level" = "level", "new_taxa" = "new_taxa"))
fwrite(temp, file = "~/Results/SHUVO-MILK/results/taxonomy_SHUVO-MILK_significant_otus.csv", col.names = TRUE, sep = ",")
print (dd)

# temp$new_taxa <- paste(temp$new_taxa,", Timepoint:", temp$timepoint)
# temp <- temp %>% filter(level == "genus")
```

```{r, echo=FALSE, warning=FALSE}
load("~/Results/SHUVO-MILK/results/taxonomy_SHUVO-MILK_tot.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

D0 <- mutate(D0, avg_counts = avg_counts+1) %>% spread(key = sample_type, value = avg_counts)

D1 <- DX %>%
  inner_join(D0, by = c("level" = "level", "new_taxa" = "new_taxa")) %>%
  mutate(p.value = -log10(p.value)) %>%
  gather(key = "sample_type", value = "counts", -c(level,new_taxa, p.value))

D1$level <- factor(D1$level, levels = c("genus")) #, "species"

D1 <- D1 %>% group_by(level) %>% mutate(tot = sum(counts), relab = counts/tot)

p <- ggplot(D1, aes(x = sample_type, y = new_taxa))
p <- p + geom_tile(aes(fill = relab), colour = "white")
p <- p + theme(legend.position="bottom")
p <- p + facet_grid(level~sample_type, scales = "free",  space = "free")
p <- p + xlab("sample_types") + ylab("Taxa")
p <- p + scale_fill_gradient(low = "orange", high = "blue")
p <- p + theme(strip.text.y = element_text(size = 5), 
               strip.text.x = element_text(size = 6),
               # axis.text.y = element_text(size = 4),
               axis.text.x = element_text(size = 6),
               axis.title = element_text(size = 6))
p <- p + guides(fill="none") + theme(axis.title.y = element_blank(),
                                     axis.text.y = element_blank(),
                                     axis.ticks.y = element_blank())
p

dd <- D1 %>% mutate(variable = "p-value")

q <- ggplot(dd, aes(x = factor(1), y = new_taxa, group=level))
q <- q + geom_tile(aes(fill = p.value), colour = "white")
q <- q + theme(legend.position="bottom")
q <- q + facet_grid(level~variable, space="free", scales = "free_y")
q <- q + scale_fill_gradient(low = "orange", high = "blue")
q <- q + xlab("sample_types") + ylab("Taxa")
q <- q + theme(strip.text = element_text(size = 4), 
               strip.text.x = element_text(size = 6),
               axis.text.y = element_text(size = 6),
               axis.title = element_text(size = 6))
q <- q + guides(fill=FALSE) + theme(
  # axis.title.x = element_blank(),
  # axis.text.x=element_blank(),
  # axis.ticks.x=element_blank(),
  strip.text.y = element_blank(),
  # axis.text.x = element_blank()
  axis.text.x = element_text(size = 6)
)
q <- q + xlab("") 
q
#ggsave(filename = "~/Results/SHUVO-MILK/results/taxonomy_heatmap_pvalue_SHUVO-MILK.png", plot = q, device = "png", width = 6, height = 6)


figure_final <- ggarrange(q, p, widths=c(0.1, 0.1), labels=c("A", "B"), legend = "bottom", common.legend = TRUE)

print(figure_final)
ggsave(filename = "~/Results/SHUVO-MILK/results/taxonomy_heatmap_SHUVO-MILK.png", plot = figure_final, device = "png", width = 7, height = 10)
```

<!-- ### Extracting estimated coefficients for "significant" OTUs -->

<!-- ```{r} -->
<!-- res = data.frame("level"=NULL, "OTU"=NULL, "p_value"=NULL, "Control"=NULL,"Treated"=NULL) -->

<!-- DX <- DX %>% filter(level == "genus") -->
<!-- m1 <- m1 %>% filter(level=="genus") -->
<!-- # DX$new_taxa <- paste(DX$new_taxa,", Timepoint:", DX$timepoint) -->
<!-- # m1$new_taxa <- paste(m1$new_taxa,", Timepoint:", m1$timepoint) -->


<!-- for (name in DX$new_taxa) { -->

<!--   print(paste("analysing OTU ", name)) -->
<!--   pval = as.numeric(DX[DX$new_taxa==name,"p.value"]) -->
<!--   level = DX[DX$new_taxa==name,"level"] -->

<!--   ## estimating coefficients -->
<!--   temp2 = filter(m1, new_taxa == name) -->
<!--   temp2$sample_type <- factor(temp2$sample_type, levels = c("control","EO")) -->
<!--   g = lm(counts ~ sample_type, data = temp2) -->

<!--   ## extracting coefficients -->
<!--   coefs = g$coefficients -->
<!--   coefs = coefs[!grepl("(Intercept)", names(coefs))] -->
<!--   names(coefs) = gsub("sample_type","",names(coefs)) -->
<!--   coefs = as.data.frame(t(coefs)) -->

<!--   ## adding metadata -->
<!--   coefs["level"] = level -->
<!--   coefs["OTU"] = name -->
<!--   coefs["p_value"] = pval -->

<!--   # saving results -->
<!--   res = rbind.data.frame(res, coefs) -->
<!-- } -->

<!-- write.csv(res, "~/Results/SHUVO-MILK/results/taxonomy_res.csv") -->
<!-- ``` -->

```{r}
Dna <- filter(temp, level %in% c("genus"))
Dna$level <- NULL
Dna <- as.data.frame(t(Dna))
colnames(Dna) <- Dna[1,]
Dna <- Dna [3:5,]

names <- rownames(Dna)
rownames(Dna)<- NULL
Dna <- cbind(names, Dna)

# Dna$`Healthy quarter` <-NULL
# Dna$`Subclinical quarter` <-NULL

# write.csv(Dna, "~/Results/SHUVO-MILK/results/taxonomy_res3.csv")

q <- ggplot(Dna, aes(x = factor(1), y = new_taxa, height=0.95))
q <- q + geom_tile(aes(fill = p.value), colour = "white")
q <- q + scale_fill_gradient(low = "red", high = "green", limits=c(0,0.05)) #, guide = "none"
q <- q + xlab("pvalue") + ylab("Genera")
q <- q + theme(axis.text.x = element_text(size = 13),
               axis.text.y = element_text(size = 13),
             axis.ticks = element_blank(), 
             legend.position = "bottom",
             axis.title.x = element_blank(),
             axis.title.y = element_text(size=14))
q

fg <- ggplot(Dna, aes(x= factor(1), y=names)) 
# , y=`Healthy quarter`)) + geom_bar(width=0.95,stat="identity")
# fg <- fg + coord_flip() 
fg <- fg + ylab("Behaviour Treated vs. Control") + xlab("Taxa")
fg <- fg + theme(axis.text.x = element_text(size = 11),
                 axis.text.y = element_blank(),
                 axis.ticks = element_blank(),
                 legend.position = "bottom",
                 axis.title.x = element_text(size=14),
               axis.title.y = element_blank())
fg
 
# figure_final <- ggarrange( q2, fg2, q3, fg3, widths=c(1,1,1,1), heights=c(1,1,1,1), labels = c("A","B","C","D"), legend = "bottom")
# print(figure_final)

#ggsave(filename = "~/Results/SHUVO-MILK/results/taxonomy_compl_genera_T23.png", plot = figure_final, device = "png", width = 16, height = 6)
```
