---
title: "Taxonomy_Skinswabs"
author: "Chiara Gini"
date: "24/05/2022"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library("DT")
library("knitr")
library("broom")
library("vegan")
library("ggpubr")
library("tibble")
library("janitor")
library("ggplot2")
library("ggrepel")
library("cowplot")
library("reshape2")
library("tidyverse")
library("data.table")
```

# CURRENTLY MISSING:

- boxplots with avg_abundances corrected for baseline
- barplots for all the taxa combined together

# COMPARISON BETWEEN TREATMENTS

## SET UP

#### reading the metadata

```{r, echo=FALSE}
# basedir="~/Documents/MILKQUA"
# prjdir = "Analysis/milkqua_skinswab/qiime1.9"
# outdir = "results"
# 
# fname = file.path(basedir,"Config/manifest_skinswab.csv")
# metadata <- fread(fname)

basedir="~/Results"
prjdir = "SKINSWABS"
outdir = "results"

fname = file.path(basedir, prjdir, "mapping_milkqua_skinswabs.csv")
metadata <- fread(fname)
names(metadata)[1] <- "sample"

newt = c("T1","T2","T3")
oldt = unique(metadata$timepoint)
metadata$timepoint_recoded = newt[match(metadata$timepoint, oldt)]
metadata <- rename(metadata, timepoint_orig = timepoint, timepoint = timepoint_recoded)

metadata %>%
  group_by(timepoint,treatment) %>%
  dplyr::summarise(N=n())

metadata$treatment <- toupper(metadata$treatment)
```

#### reading the normalised OTU table

- !! **sample n. 29 has been removed** when filtering for sequencing quality !!

```{r, echo=FALSE}
fname = file.path(basedir, prjdir, outdir, "otu_norm_CSS.csv")
otu <- fread(fname)
```

## Core microbiota

We want to identify the OTU (at any given taxonomic level) that are shared by the majority of the samples. To this aim, we set a threshold (e.g. $95\% = 0.95$) and proceed to calculate the proportion of samples where each OTU (e.g. each Genus, each Genus etc.) is not absent:

- we group by `OTU` (e.g. Genus) **and** `sample`
- we sum the corresponding counts (e.g. all OTUs belonging to each Genus in any sample)
- we count how many such sums are zero/non-zero and take the ratio over the total n. of samples
- we filter based on this ratio (non-zero samples / all samples) and the desired threshold

**NOTE: the following scripts fits perfectly for every taxa level, but only when singularly run: you will need to run every level separately, by changing the name of the taxa inside the scripts. It can be easily done via Find+Replace function in RStudio (search symbol in the bar above). Careful: in some cases levels are empty: this is currently a problem to be solved.**
 
```{r taxa setup, echo =FALSE}
threshold = 0.99
```

```{r core_micro, echo =FALSE}
temp = select(otu, contains("sample-"), Genus) %>% gather(key = "sample", value = "counts", -Genus)
core = temp %>% group_by(Genus, sample) %>%
  dplyr::summarise(tot = sum(counts)) %>%
  group_by(Genus) %>%
  dplyr::summarise(N=n(), zeros = sum(tot == 0), ratio = (1-zeros/N)) %>%
  filter(ratio >= threshold)


temp = filter(temp, Genus %in% core$Genus)
```

### Genus core microbiota

Adding metadata to the Genus core microbiota data:

```{r Genus average abundances, echo=FALSE}
temp$timepoint <- metadata$timepoint[match(temp$sample, metadata$sample)]
temp$treatment <- metadata$treatment[match(temp$sample, metadata$sample)]
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
x = temp %>%
  group_by(timepoint) %>%
  dplyr::mutate(tot = sum(counts)) %>%
  dplyr::group_by(timepoint, Genus) %>%
  dplyr::summarise(phyltot = sum(counts), proportion = round(phyltot/tot, 3)) %>%
  unique()
```

Make tables and plots of the Genus core microbiota over time: 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
x %>%
  group_by(timepoint, Genus) %>%
  summarise(avg = mean(proportion)) %>%
  spread(key = timepoint, value = avg) %>%
   adorn_totals("row") 
```

```{r, fig.height=8, fig.width=10}
p <- ggplot(x, aes(x=factor(1), y=proportion, fill=Genus)) + geom_bar(width=1,stat="identity")
p <- p + coord_polar(theta='y') #activate this if you want a piechart
p <- p + geom_label(aes(label = 100*proportion), show.legend = FALSE, size = 3)
# p <- p + geom_label_repel(aes(label = paste0(proportion*100, "%")), position=position_stack (vjust=0.4), size=3, show.legend = F, force = 0.6,  max.overlaps = 10)
p <- p + facet_wrap(~timepoint)
# p <- p + scale_fill_brewer(palette = "Set1") 
p <- p + guides(fill = guide_legend(title = "Genus")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
p <- p + xlab("Relative abundances") + ylab("Percentages")
p <- p + theme(text = element_text(size=8),
               strip.text = element_text(size = 8),
               axis.text.y = element_blank(),
               axis.text.x = element_blank(),
               axis.ticks = element_blank(), 
               axis.ticks.y = element_blank(),
               axis.ticks.x = element_blank(),
               legend.text=element_text(size=6),
               legend.title=element_text(size=6))
p

#ggsave(filename = file.path(basedir, prjdir, outdir, "taxonomy_Genus_piechart_skinswabs_wbaseline.png"), plot = p, device = "png", width = 14, height = 7)
```

#### Core microbiota: family, Genus etc. continue if needed = replace Genus with Genus or whatevere you are interested in. Careful: blank spaces can suggest that it was not possible to identify that taxa level. --> solve this problem?

## Correction for baseline

In our case it is not practical to follow the same subject (animal) from T1 to T3 (some animals have data on multiple quarters); therefore we must adopt a correction strategy based on average counts at baseline:

- first, sum all counts by OTU level (e.g. many OTUs belonging to the same Genus)
- then calculate average counts at baseline (together for treated and controls: they're supposed to be equivalent at T1 --> no treatment yet !)
- finally, subtract average at baseline from all counts: in this way, the average at T1 will be zero for both treatments and controls, while averages at later timepoints will reflect relative shifts in the experimental groups
- corrected counts can be rescaled to be more easily comparable

```{r}
temp = select(otu, contains("sample"), Genus) %>% 
  gather(key = "sample", value = "counts", -Genus) %>%
  group_by(Genus, sample) %>%
  summarise(tot = sum(counts))

temp$timepoint <- metadata$timepoint[match(temp$sample, metadata$sample)]
temp$treatment <- metadata$treatment[match(temp$sample, metadata$sample)]
```

```{r}
bl = filter(temp, timepoint == "T1") %>% 
  group_by(Genus, treatment) %>%
  summarise(bl = mean(tot))
```

```{r}
temp <- temp %>% 
  inner_join(bl, by = c("Genus" = "Genus", "treatment" = "treatment")) %>%
  mutate(corrected_counts = tot - bl)
```


```{r}
temp <- temp %>%
  group_by(Genus) %>%
  mutate(scaled_counts=scales::rescale(corrected_counts,c(0,100)))
```

#### Sanity check plot

Is baseline really at 0?

```{r}
dd <- temp %>%
  group_by(Genus, timepoint, treatment) %>%
  summarise(avg = mean(corrected_counts))

p <- ggplot(data=dd, aes(x=timepoint, y=avg, group=treatment)) +
  geom_line(aes(colour=treatment)) +
  geom_point()
p <- p + facet_wrap(~Genus, scales = "free") 
p

#ggsave(filename = file.path(basedir, prjdir, outdir, "taxonomy_Genus_baseline_check.png"), plot = p, device = "png", width = 20, height = 20)
```

#### Additional plots and tables (e.g. bubble chart etc.) if desired (@chiara)

## Effect of treatment

The significance of differences between treatments (oil vs non-oil) was evaluated using a linear model (ANOVA) that included the effect of timepoint and treatments:

$$
counts_{ikjt} = \mu + timepoint_{k(j)} + treatment_{t(j)} + e_{ikjt}
$$

this way, the variability due to region is removed when assessing the effect of treatments.

```{r lm, echo=FALSE, warning=FALSE, message=FALSE}
library("lmerTest")

mm <- filter(temp, timepoint != "T1")

D <- mm %>%
  group_by(Genus) %>%
  do(tidy(anova(lm(corrected_counts ~ timepoint + treatment, data=.)))) %>%
  filter(term == "treatment")

datatable(D, options = list(pageLength=100)) %>% 
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))

fname = file.path(basedir, prjdir, outdir, "significant_taxa.csv")
filter(D, p.value <= 0.05) %>% select(c(Genus,term,p.value)) %>% fwrite(file = fname, sep = ",", col.names = TRUE)
```

```{r, echo=FALSE}
dd <- temp %>%
  group_by(Genus,timepoint,treatment) %>%
  summarise(avg = mean(tot)) %>%
  spread(key = treatment, value = avg)

d0 <- select(D, c(Genus, p.value))

dd <- dd %>% inner_join(d0, by = c("Genus" = "Genus"))

to_save = list(D,dd,d0)
fname = file.path(basedir, prjdir, outdir, "taxonomy_group.RData")
save(to_save, file = fname)
```

```{r, echo=FALSE, warning=FALSE}
load(fname)
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

Dna <- DX[!(is.na(DX$Genus) | DX$Genus==""), ]
Dna <- filter(Dna, p.value < 0.05)

## remember here to comment the taxonomic level you are using
Dna$Phylum <- NULL
Dna$Kingdom <- NULL
Dna$Species <- NULL
Dna$Order <- NULL
Dna$Family <- NULL
#Dna$Genus <- NULL
Dna$Class<- NULL
Dna$ControlvsTreated <- (Dna$CONTROL-Dna$TREATED)
Dna$CONTROL <-NULL
Dna$TREATED <-NULL

p <- ggplot(Dna, aes(x=p.value, y=Genus, fill=p.value))
p <- p + geom_bar(width=1,stat="identity", position = "fill")
p <- p + scale_fill_gradient(low = "orange", high = "blue")
p <- p + xlab("pvalue") + ylab("Genus")
p

#ggsave(filename = file.path(basedir, prjdir, outdir, "barplot_pvalues2.png"), plot = p, device = "png", width = 8, height = 15)
```

```{r}
fg <- ggplot(Dna, aes(x=Genus, y=ControlvsTreated, fill=timepoint)) + geom_bar(width=1,stat="identity")
fg <- fg + coord_flip()
fg <- fg + ylab("Behaviour of the Treated against Control, in significant p_values") + xlab("Taxa")
fg <- fg + theme(axis.text.x = element_text(size = 6),
               axis.title = element_text(size = 6))
fg

# ggsave(filename = file.path(basedir, prjdir, outdir, "barplot_significantOTUsbehaviour.png"), plot = fg, device = "png", width = 15, height = 40)
```

```{r}
figure_final <- ggarrange(p, fg, widths=c(1, 1), labels=c("A", "B"), legend = "bottom")

print(figure_final)
#ggsave(filename = "~/Results/SKINSWABS/results/taxonomy_heat+barplot_skinswabs.png", plot = dragon, device = "png", width = 7, height = 10)
```
