---
title: "taxonomy per treatment and timepoint: milk samples"
author: "Filippo Biscarini"
date: "November 19, 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("DT")
library("knitr")
library("dplyr")
library("tidyr")
library("broom")
library("vegan")
library("ggpubr")
library("ggplot2")
library("reshape2")
library("data.table")
```

## 16S sequencing data - cow milk samples

```{r data, echo=FALSE}
metadata <- fread("mapping_file.csv")
names(metadata)[1] <- "sample"
metadata = filter(metadata, sample_type == "milk")
metadata %>%
  group_by(timepoint,treatment,cow) %>%
  dplyr::summarise(N=n()) %>%
  spread(key = "timepoint", value = N)
```

```{r taxonomy, echo=FALSE}
otu <- read.table("results/taxa_summary_abs/mapping_file_L2.txt",header=TRUE, sep="\t" ,comment.char = "&")
names(otu)[1] <- "sample"

otu <- otu %>%
  select(-c(index,index2,I7_Index_ID,I5_Index_ID,extraction_serial)) %>%
  filter(sample_type == "milk")

otu$treatment = metadata$treatment[match(otu$Sample_ID,metadata$Sample_ID)]
otu$timepoint = metadata$timepoint[match(otu$Sample_ID,metadata$Sample_ID)]
otu$timepoint = factor(otu$timepoint, levels = c("0","8","21","28"))

## relative abundances
metadata_cols = seq(1,7)
M <- otu[,-metadata_cols]
M <- M/rowSums(M)
M <- bind_cols(otu[,metadata_cols],M)
```


```{r, echo=FALSE, fig.height=12, fig.width=12}
metadata_cols = names(otu)[1:7]
mO <- reshape2::melt(otu,id.vars = metadata_cols, value.name = "counts", variable.name = "phylum")
mO$phylum <- gsub("^.*\\.","",mO$phylum)
mO[mO$phylum == "2","phylum"] <- "WPS.2"
mO$sample <- factor(mO$sample)
mO <- mO %>%
  arrange(treatment)
```

```{r, echo=FALSE, fig.height=12, fig.width=12}
m1 <- reshape2::melt(M,id.vars = metadata_cols, value.name = "counts", variable.name = "phylum")
m1$phylum <- gsub("^.*\\.","",m1$phylum)
m1[m1$phylum == "2","phylum"] <- "WPS.2"
m1$sample <- factor(m1$sample)
m1 <- m1 %>%
  arrange(treatment)

# m1$sample <- paste(as.character(m1$sample),as.character(m1$timepoint),sep="-")
ordered_samples <- metadata %>%
  arrange(treatment) %>%
  select(treatment,sample)

m1$sample <- factor(m1$sample, levels = ordered_samples$sample)

m1$treatment <- factor(m1$treatment)
m1$timepoint <- factor(m1$timepoint)
```


### Treatments and timepoints

First, let's look at phyla:

```{r, echo=FALSE, warning=FALSE, message=FALSE}
D <- m1 %>%
  group_by(phylum,treatment,timepoint) %>%
  dplyr::summarise(N=n(), avg_abund = round(mean(counts),4), std = round(sd(counts),3)) %>%
  filter(avg_abund>0.005) %>%
  arrange(phylum,avg_abund)

phylum <- D %>%
  group_by(phylum) %>%
  dplyr::summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  select(phylum,s)

D$phylum <- factor(D$phylum, levels = rev(phylum$phylum[order(phylum$s)]))
D <- D %>%
  arrange(treatment,phylum)

kable(D)
```


```{r, echo=FALSE, fig.height=8, fig.width=12}
p <- ggplot(mO, aes(x=phylum,y=as.factor(cow),fill=counts))
p <- p + geom_tile()
p <- p + facet_grid(treatment~timepoint, scales = "free")
p <- p + ylab("sample")
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1),axis.text.y = element_text(size=9))
p <- p + scale_fill_gradient2(low="ivory",high = "red")
p
```

```{r, echo=FALSE}
my_palette = get_palette(c("#00AFBB", "#E7B800", "#FC4E07","darkgrey","beige"), length(unique(D$phylum)))

p <- ggplot(D, aes(x=factor(1), y=avg_abund, fill=phylum)) + geom_bar(width=1,stat="identity")
p <- p + facet_grid(timepoint~treatment)
p <- p + coord_polar(theta='y')
p <- p + xlab("relative abundances") + ylab("")
p <- p + scale_fill_manual(values = my_palette)
p <- p + theme(text = element_text(size=15),
               axis.text.x = element_text(size=11),
               # axis.text.y = element_text(size=12),
               strip.text = element_text(size = 16),
               axis.text.y = element_blank(),
               axis.ticks.y = element_blank(),
               legend.text=element_text(size=10),
               legend.title=element_text(size=10))

print(p)
ggsave(filename = "figures/phylum_piechart.png", plot = p, device = "png")
```


### Treatments, all taxonomic levels

```{r,echo=FALSE,message=FALSE,warning=FALSE}
D <- D %>%
  group_by(treatment) %>%
  filter(avg_abund > 0.01) %>%
  arrange(desc(avg_abund))

phylum <- D %>%
  group_by(phylum) %>%
  dplyr::summarize("s"=sum(avg_abund)) %>%
  arrange(desc(s)) %>%
  select(phylum,s)

D$phylum <- factor(D$phylum, levels = rev(phylum$phylum[order(phylum$s)]))
D <- D %>%
  arrange(phylum)

D$taxa_level = rep("phylum", nrow(D))

otu <- fread("results/taxa_summary_abs/mapping_file_L6.txt",header=TRUE)
names(otu)[1] <- "sample"

otu <- otu %>%
  select(-c(index,index2,I7_Index_ID,I5_Index_ID,extraction_serial)) %>%
  filter(sample_type == "milk")

otu$treatment = metadata$treatment[match(otu$Sample_ID,metadata$Sample_ID)]
otu$timepoint = metadata$timepoint[match(otu$Sample_ID,metadata$Sample_ID)]
otu$timepoint = factor(otu$timepoint, levels = c("0","8","21","28"))

mO <- reshape2::melt(otu,id.vars = metadata_cols, value.name = "counts", variable.name = "taxa")
mO$level <- sapply(mO$taxa, function(x) length(strsplit(as.character(x),";")[[1]]) ) 

get_taxa <- function(x,n) {
  
  z <- strsplit(x, split = ";")
  res = lapply(z, function(x) x[n])
  return(unlist(res))
}

new_taxa <- c(NULL)
for (i in seq_along(mO$counts)) {
  
  arg1 = as.character(mO$taxa[i])
  arg2 = ifelse(mO$level[i] <= 6, mO$level[i], 6)
  
  rr = get_taxa(arg1,arg2)
  new_taxa <- c(new_taxa,rr)
}

mO$new_taxa <- new_taxa

table(mO$level)
neues <- c("phylum","class","order","family","genus")
altes <- seq(2,6)
mO$level <- neues[match(mO$level,altes)]

## relative abundances
mO <- mO %>%
  dplyr::group_by(level,sample) %>%
  dplyr::mutate(tot = sum(counts), rel_abundance = counts/tot)


D <- mO %>%
  dplyr::group_by(level,new_taxa,treatment,timepoint) %>%
  dplyr::summarise(avg_abund = round(mean(rel_abundance),4), std = round(sd(rel_abundance),3))
```

A table of relative abundances (and standard deviations) per taxa and treatment:

```{r, tab2, echo=FALSE}
D <- D %>%
  dplyr::group_by(new_taxa,timepoint,treatment) %>%
  dplyr::filter(avg_abund > 0.01) %>%
  arrange(level,desc(avg_abund))

kable(D)
```

Same results, but in wide-format (treatments on the horizontal axis): relative abundances (no standard deviations):


```{r, echo=FALSE}
D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))

D0 <- D %>%
  select(-c(std)) %>%
  spread(treatment,avg_abund)

# D0[is.na(D0)] <- ""

bg.picker <- function(z) {
  
  if(is.na(z)) {return(NULL)}
  if(is.numeric(z)) {
    
    if(z > 0.05 & z <= 0.10) {return("pink")}
    if(z > 0.10) {return("yellow")}
  }
}

library("formattable")
formattable(D0, list(
  
  `Control` = color_tile("white","green"),
  `Treated` = color_tile("white","deepskyblue")
  # `3 PR` = color_tile("white","yellow"),
  # `4 PR` = color_tile("white","pink"),
  # `5 PR` = color_tile("white","yellow"),
  # `6 PR` = formatter("span",
  #                 style = x ~ style(
  #                   "border-radius" = "4px",
  #                   "padding-right" = "4px",
  #                   "background-color" = sapply(x, bg.picker))
  #                 )
))

```

**timepoint**

```{r, echo=FALSE}
D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))

D0 <- D %>%
  select(-c(std)) %>%
  spread(timepoint,avg_abund)

# D0[is.na(D0)] <- ""

bg.picker <- function(z) {
  
  if(is.na(z)) {return(NULL)}
  if(is.numeric(z)) {
    
    if(z > 0.05 & z <= 0.10) {return("pink")}
    if(z > 0.10) {return("yellow")}
  }
}

library("formattable")
formattable(D0, list(
  
  `0` = color_tile("white","yellow"),
  `8` = color_tile("white","pink"),
  `21` = color_tile("white","yellow"),
  `28` = color_tile("white","pink")
  # `5 PR` = color_tile("white","yellow"),
  # `6 PR` = formatter("span",
  #                 style = x ~ style(
  #                   "border-radius" = "4px",
  #                   "padding-right" = "4px",
  #                   "background-color" = sapply(x, bg.picker))
  #                 )
))

```

Finally,a bubble chart to visualize relative abundances per taxa and treatment:

```{r, echo=FALSE, fig.height = 14, fig.width = 10}
taxa <- D %>%
  dplyr::group_by(level,new_taxa,timepoint) %>%
  dplyr::summarize("s"=sum(avg_abund)) %>%
  arrange(desc(level),s) %>%
  select(level,new_taxa,s)

taxa <- taxa %>%
  dplyr::group_by(level) %>%
  arrange(level,s)

# D$taxa <- factor(D$taxa, levels = rev(taxa$taxa[order(taxa$s)]))
# 
# D <- D %>%
#   arrange(level,taxa)

p <- ggplot(D, aes(x = treatment, y = new_taxa))
p <- p + geom_point(aes(size = avg_abund, colour = treatment), alpha = 0.4)
p <- p + facet_grid(level~timepoint, scales = "free", space = "free")
p <- p + scale_color_manual(values = c("#00AFBB", "#E7B800", "#FC4E07"))
p <- p + theme(axis.text.x = element_text(angle = 90, hjust = 1))
p <- p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
p <- p + theme(axis.text.x = element_text(size=8))
p <- p + theme_bw()
p
```

### Significance of treatments (+ timepoint)

The significance of differences between treatments (milk vs waste-milk) was evaluated using a linear model (ANOVA) that included the effect of region (rumen, ileum) and treatments:

$$
counts_{ikj} = \mu + timepoint_k + treatment_j + e_{ikj}
$$

this way, the variability due to region is removed when assessing the effect of treatments.

```{r lm, echo=FALSE}
m1 <- mO %>%
  filter(!is.na(level)) %>%
  arrange(level,treatment)

D <- m1 %>%
  dplyr::group_by(level, new_taxa) %>%
  do(tidy(anova(lm(counts ~ timepoint+treatment, data = .)))) %>%
  filter(term == "treatment")

D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>% 
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))

```


### Significance of treatments (per timepoint)

The significance of differences between treatments (milk vs waste-milk) was evaluated using a linear model (ANOVA) that included the effect of region (rumen, ileum) and treatments:

$$
counts_{ij} = \mu + treatment_j + e_{ij}
$$

this way, the variability due to region is removed when assessing the effect of treatments.

```{r lm_2, echo=FALSE}
m1 <- mO %>%
  arrange(level,treatment)

dd_counts <- mO %>%
  dplyr::group_by(level,new_taxa,timepoint,treatment) %>%
  # summarise(avg = mean(counts)) %>%
  dplyr::summarise(avg = mean(rel_abundance)) %>%
  spread(key = "treatment", value = "avg")

group_by(dd_counts, level) %>% summarise(tot = sum(Treated))

D <- m1 %>%
  dplyr::group_by(level, new_taxa, timepoint) %>%
  do(tidy(anova(lm(counts ~ treatment, data = .)))) %>%
  filter(term == "treatment")

D$level  <- factor(D$level,levels = c("phylum","class","order","family","genus"))
D <- D %>%
  arrange(level,new_taxa)

datatable(D, options = list(pageLength=100)) %>% 
  formatStyle('p.value', backgroundColor = styleInterval(0.05, c('yellow', 'white')))

```


```{r}
DX <- D %>%
  filter(`p.value` <= 0.05) %>%
  select(c(level,new_taxa,timepoint,`p.value`)) %>%
  arrange(level,`p.value`)

D0 <- mO %>%
  dplyr::group_by(level,new_taxa,timepoint,treatment) %>%
  dplyr::summarise(avg_counts = mean(counts))

to_save = list(D,DX,D0)
save(to_save, file = "taxonomy_timepoint.RData")

load("taxonomy_timepoint.RData")
D <- to_save[[1]]
DX <- to_save[[2]]
D0 <- to_save[[3]]

dd <- spread(D0, key = treatment, value = avg_counts)
temp <- inner_join(DX,dd, by = c("level" = "level", "new_taxa" = "new_taxa", "timepoint" = "timepoint"))
fwrite(temp, file = "significant_otus.csv", col.names = TRUE, sep = ",")
```

